---
title: "6: Preparation of LTA Data Mall Data for Take-home Exercise 2"
author: "Emma Chew En Chin"
date: "Oct 4, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## Loading the R packages

```{r}
pacman::p_load(tidyverse, sf, sfdep, tmap, knitr, httr, rvest)
```

## Importing Data

### Bus Stop Location

As Bus Stop location data has crs = 9001, we standardise it into crs = 3414.

```{r}
BusStop = st_read(dsn = "data/LTADataMall/",
                  layer = "BusStop") %>%
  st_transform(crs = 3414)
```

### Master Plan 2019 Subzone

```{r}
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>%
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
mpsz_sf <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS", 
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")
```

```{r}
write_rds(mpsz_cl, 
          "data/rds/mpsz_cl.rds")
```

### O-D Bus Trips

```{r}
odbus <- read_csv("data/LTADataMall/origin_destination_bus_202508.csv")
```

## Visualise the Busstops

We see some busstops outside the polygons. (Due to JB border)

```{r}
tm_shape(mpsz_cl) + 
  tm_polygons() + 
  tm_shape(BusStop) + 
  tm_dots()
```

### Extract Bus Stops within Singapore

We do not use left join as we only want to get bus stops within mpsz.

```{r}
BusStop_in_SG <- st_join(
  BusStop, mpsz_cl,
  join = st_within,
  left = FALSE)
```

Plot the bus stop data again to check that the bus stops have been removed.

```{r}
tm_shape(mpsz_cl) + 
  tm_polygons() + 
  tm_shape(BusStop_in_SG) + 
  tm_dots()
```

## Deriving Analytical Hexagon

```{r}
hexagon <- st_make_grid(mpsz_cl,
                        cellsize = 700,
                        what = "polygon",
                        square = FALSE) %>%
  st_sf()
```

### Filtering hexagons with bus stops

`n_collisions` counts the number of bus stops in each hexagon.
Only retain hexagons with bus stops.

```{r}
hexagon$n_collisions =
  lengths(st_intersects(
    hexagon, BusStop_in_SG))

hexagon <- filter(hexagon, n_collisions > 0)
```

### Assigning ids to each hexagon

Create id for each hexagon as previous hexagon table has no id.

```{r}
hexagon <- hexagon %>%
  select(,-n_collisions)

hexagon$HEX_ID <- sprintf(
  "H%04d", seq_len(nrow(hexagon))) %>%
  as.factor()

kable(head(hexagon))
```

## Preparing Trip Generation Data

Only interested in ORIGIN_PT_CODE (trip origin).

### Cleaning the data

Remove unnecessary columns.

```{r}
trips <- odbus %>%
  select(c(ORIGIN_PT_CODE, DAY_TYPE, TIME_PER_HOUR, TOTAL_TRIPS)) %>%
  rename(BUS_STOP_N = ORIGIN_PT_CODE) %>%
  rename(HOUR_OF_DAY = TIME_PER_HOUR) %>%
  rename(TRIPS = TOTAL_TRIPS)
trips$BUS_STOP_N <- as.factor(trips$BUS_STOP_N)
kable(head(trips))
```

### Combining bus stop and hexagon layers

```{r}
bs_hex <-st_intersection(
  BusStop, hexagon) %>%
  st_drop_geometry() %>%
  select(c(BUS_STOP_N, HEX_ID))
kable(head(bs_hex))
```

### Adding Hex_ID into bus trips data

```{r}
trips <- inner_join(trips, bs_hex)
kable(head(trips))
```
