---
title: "10: Market Area Analysis"
author: "Emma Chew En Chin"
date: "Nov 15, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Installing R packages

```{r}
pacman::p_load(sf, tidyverse, MCI, tmap, ggstatsplot)
```

## 2 Importing Geospatial Data

```{r}
stores <- st_read(
  dsn = "data/geospatial",
  layer = "STORES_TC")
```

```{r}
town <- st_read(
  dsn = "data/geospatial",
  layer = "TOWN_TC")
```

```{r}
village <- st_read(
  dsn = "data/geospatial",
  layer = "VILLAGE_TC")
```

The output print above shown that the three data sets are in TWD97 / TM2 zone 121 (i.e epsg: 3826) projected coordinates systems.

```{r}
tmap_mode("view")
tm_shape(town) +
  tm_polygons() +
tm_shape(village) +
  tm_polygons() + 
tm_shape(stores) +
  tm_dots()
tmap_mode("plot") 
```

## 3 Importing Aspatial Data

```{r}
pop2019 <- read_csv(
  "data/aspatial/pop2019.csv")

store_sales <- read_csv(
  "data/aspatial/Store_sales_data.csv")

```

```{r}
summary(store_sales)
```

## 4 Data Preparation

### 4.1 Aggregating store sales

Aggregate store sales data at Store code

```{r}
store_sales_aggre <- store_sales %>%
  group_by(`Store Code`) %>%
  summarise(
    "Total_Bill" = sum(Bills, na.rm = TRUE),
    "Total_Sales" = sum(Amount, na.rm = TRUE))
```

### 4.2 Relational join

`left_join()` of **dplyr** package is used to append data from *store_sales* data.frame onto stores sf data.frame. At the same time a new field called *Store_size* has been created.

```{r}
stores <- stores %>%
  left_join(store_sales_aggre,
            by = join_by(
              Store_CD == `Store Code`)) %>%
  mutate(Store_size = 100)
```

Next, `summary()` is used to check if the are any missing values in the data.frame.

```{r}
summary(stores)
```

### 4.3 Tidying pop2019 data.frame

Before joining pop2019 and village sf dataframes, we will check that the values in the join field are the same.

```{r}
glimpse(pop2019)
```

```{r}
glimpse(village)
```

Village ID in pop2019 dataframe is not in same structure as village code in village dataframe.

Hence we will remove the “-” sign from the values.

```{r}
pop2019 <- pop2019 %>%
  mutate(V_ID = str_remove_all(
    V_ID, "-"))
```

pop2019 dataframe is then joined onto village sf dataframe.

```{r}
village <- village %>%
  left_join(pop2019,
            by = join_by(
              VILLCODE == V_ID))
```

### 4.4 Checking for missing or zero values

This is done using summary.

```{r}
summary(village)
```

The output print above revealed that the are five missing values in H_CNT, P_CNT, M_CNT and F_CNT fields.

Hence, code chunk below is used to remove the five records with missing values from village sf data.frame.

```{r}
village <- village %>%
  filter(!is.na(P_CNT)) 
```

```{r}
any(is.na(village$H_CNT))
```

```{r}
any(is.na(village$P_CNT))
```

```{r}
any(is.na(village$M_CNT))
```

```{r}
any(is.na(village$F_CNT))
```

## 5 Computing Distance Matrix

Instead of calculating the distance between the villages and stores, we will calculate the centroid of each village using the code chunk below. Then, compute the distance between the centroids of the villages and the stores.

```{r}
village_center <- st_point_on_surface(village)
distmat <- st_distance(village_center, stores) 
```

Next, the code chunk below is used to convert the distance matrix into long data.frame format

```{r}
rownames(distmat) <- village$`VILLCODE`
colnames(distmat) <- stores$`Store_CD`

distmat_long <- distmat %>%
  as.data.frame() %>%
  rownames_to_column(var = "VILLCODE") %>%
  pivot_longer(
    cols = -VILLCODE,
    names_to = "Store_CD",
    values_to = "distance"
  ) %>%
  mutate(distance = as.numeric(distance)) 
```

## 6 Preparing Interaction Matrix

Interaction matrix required by MCI package is built.

```{r}
interaction_matrix <- distmat_long %>%
  left_join(
    stores %>% 
      st_drop_geometry() %>% 
      select(Store_CD, Store_size),
    by = "Store_CD") %>%
  relocate(
    distance, .after = last_col())
```

```{r}
summary(interaction_matrix)
```

## 7 Computing Huff Model

```{r}
huff_share <- huff.shares(
  interaction_matrix, "VILLCODE", "Store_CD", 
  "Store_size", "distance")
```

### 7.1 Trade area analysis for a single store

First, let us select market share data of a store by using *Store_CD *field (i.e. TG).

```{r}
market_share <- huff_share %>%
  filter(Store_CD == "TG")
```

Next, trade areas of the selected will be derived by using the code chun below.

```{r}
Trade_area <- village %>%
  left_join(market_share)
```

We extract selected stores from the Stores sf dataframe.

```{r}
selected_store <- stores %>%
  filter(Store_CD == "TG")
```

We then plot the trade areas of the selected store.

```{r}
tm_shape(Trade_area) +
  tm_polygons(fill = "p_ij",
              fill.scale = tm_scale_intervals(
                style = "quantile", 
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Patronage Probability")) +
  tm_shape(selected_store) +
    tm_dots(size = 0.3,
            fill = "red",
            col = "black") +
  tm_title("Distribution of Patronage Probability of Store BT by village") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha =0.2) 
```

### 7.2 Store sales analysis

One of the usage of Huff model is to estimate the store sales by using village population.

Firstly, let us append *P_CNT* field from *village* sf.data.frame to *huff_share* data.frame.

```{r}
huff_share_pp <- huff_share %>%
  left_join(village %>% 
      st_drop_geometry() %>% 
      select(VILLCODE, P_CNT),
    by = "VILLCODE") %>%
  relocate(
    distance, .after = last_col())
```

Next, `shares.total()` of MCI package will be used to calculate the total sales.

```{r}
huff_total <- shares.total(
  huff_share_pp, "VILLCODE", "Store_CD",
  "p_ij", "P_CNT")
```

Append the estimate store sales (Ej) onto stores sf data.frame.

```{r}
store_sales_est <- stores %>%
  left_join(huff_total,
            by = join_by(
              "Store_CD" == "suppliers_single"))
```

Now, we can visualise how well the estimate sales as compare to the actual sales by using the code chunk below.

```{r}
store_sales_nongeom <- store_sales_est %>% st_drop_geometry()

ggscatterstats(
  store_sales_nongeom,
  x = sum_E_j,
  y = Total_Sales)
```

```{r}
huff_share_wider <- huff_share %>%
  select(VILLCODE, Store_CD, p_ij) %>%
    pivot_wider(names_from = Store_CD, 
                values_from = p_ij)
```