---
title: "3: Spatio-Temporal Point Patterns Analysis"
author: "Emma Chew En Chin"
date: "Sep 11, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

A spatio-temporal point process (also called space-time or spatial-temporal point process) is a random collection of points, where each point represents the time and location of an event.

Spatio-temporal point patterns analysis can be used to determine the following: - Arrival of new points in a space across time

-   Removal of existing points in a space across time
-   Movement of existing points in a space across time
-   How the entire set of points may be distributed/clustered in a space across time

## 2 Research Context

The research objectives are:

-   To find out if the locations of forest fire in Kepulauan Bangka Belitung **spatial and spatio-temporally independent**?
-   If they are not spatio-temporally independent, where and when the observed forest fire locations **tend to cluster?**

### 2.1 Installing and Loading the R packages

-   `raster` : for handling raster data (grid of pixels) in R
-   `spatstat` : performing Spatial Point Patterns Analysis such as kcross, Lcross, etc
-   `sparr` : provides functions to estimate fixed and adaptive kernel-smoothed spatial relative risk surfaces via the density-ratio method and perform subsequent inference. Fixed-bandwidth spatiotemporal density and relative risk estimation is also supported

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse)
```

### 2.2 Importing Datasets and Preparing Study Area

Import the ESRI Shapefile, and conduct coordinate projection.

To determine the EPSG code, check [EPSG.io](EPSG.io) and search for country name + zone number.

Zone number can be derived using coordinates:

-   Zone number = floor((longitude + 180) / 6) + 1
-   For longitude \~106-107°E: Zone = floor((106 + 180) / 6) + 1 = **Zone 48**
-   Hence, Kepulauan Bangka Belitung is located in **UTM Zone 48S**, which gives us EPSG code of 32748 when searching in EPSG.io.

```{r}
kbb_sf <- st_read(dsn="data/geospatial",
               layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```

### 2.3 Converting into OWIN object

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

Use class() to confirm that the output is an owin object.

```{r}
class(kbb_owin)
```

### 2.4 Importing and Preparing Forest Fire data

Forest Fire data is imported and coordinate projection is conducted.

```{r}
fire_sf <- read_csv("data/geospatial/forestfires.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"),
                       crs = 4326) %>%
  st_transform(crs = 32748)
```

As ppp object only accepts numerical or character as mark. The code chunk below is used to convert data type of acq_date to numeric.

From this conversion we can also aggregate the dates at a month and day level.

```{r}
fire_sf <- fire_sf %>% 
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, 
                           label = TRUE, 
                           abbr = FALSE)) %>%
  mutate(Week_num = week(acq_date)) %>%
  mutate(Week_fac = month(acq_date, 
                           label = TRUE, 
                           abbr = FALSE))
```

## 3 Visualising the Fire Points

A point symbol map is plotted to show the distribution of fire points.

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots()
```

### 3.1 Visualising Geographic distribution of forest fires by month

Using tmap_arrange, we can arrange the monthly maps in a 4 x 3 grid.

```{r}
#| fig-width: 12
#| fig-height: 10
#| out.width: 100%

tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots(size = 0.3) +
tm_facets(by="Month_fac", 
            free.coords=FALSE, 
            drop.units = TRUE)
```

## 4 Computing STKDE by Month

Spattemp.density() of sparr package is used to compute STKDE.

### 4.1 Extracting forest fires by month

Remove unwanted fields from fire_sf dataframe, as `as.ppp()` only needs the mark field and geometry field.

Note that as both `sf` and `raster` are loaded, the `select()` function uses `raster::select()` method, which does not work for sf objects. Hence, we indicate `dplyr::select()`.

```{r}
fire_month <- fire_sf %>% 
  dplyr::select(Month_num)
```

### 4.2 Creating ppp object

Create ppp object and check if output is in correct object class.

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

Ensure that output is numeric.

```{r}
summary(fire_month_ppp)
```

Check for any duplicated point events.

```{r}
any(duplicated(fire_month_ppp))
```

### 4.3 Including owin object

Combine ppp and owin objects.

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

Plotting ff_owin to check the correctness of an output object.

```{r}
#| fig-width: 12
#| fig-height: 10
#| out.width: 100%
plot(fire_month_owin)
```

### 4.4 Computing Spatio-temporal KDE

Using `spattemp.density()` from the sparr package STKDE can be computed.

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

From the output, we can infer that:

-   temporal value = 0.0304, showing that clusters are highly concentrated in specific months

### 4.5 Plot the Spatio-temporal KDE object

KDE is plotted for July-December 2023.

From the KDE plots, we can observe that forest fires begin in the northern area from July, and forest fires appear with greater frequency in September. It is only until end November that forest fires subside.

```{r}
#| fig-width: 12
#| fig-height: 10
#| out.width: 100%

tims <- c(7,8,9,10,11,12)
par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}
```

## 5 Computing STKDE by Day of Year

### 5.1 Creating ppp object

```{r}
fire_yday_ppp <- fire_sf %>% 
  dplyr::select(DayofYear) %>%
  as.ppp()
```

### 5.2 Including owin object

Combining both ppp and owin objects.

```{r}
fire_yday_owin <- fire_yday_ppp[kbb_owin]
summary(fire_yday_owin)
```

### 5.3 Computing Spatio-temporal KDE

```{r}
kde_yday <- spattemp.density(
  fire_yday_owin)
summary(kde_yday)
```

From the output, we can infer that:

-   temporal value = 6.3198, showing that clusters are formed across ~ 6 days, which is around a week

While we can plot maps for each day, there will be clearer temporal shifts when we look at the fire points by week.

## 6 Computing STKDE by Week

### 6.1 Extracting forest fires by week

```{r}
fire_week <- fire_sf %>% 
  dplyr::select(Week_num)
```

### 6.2 Creating ppp object

Create ppp object and check if output is in correct object class.

```{r}
fire_week_ppp <- as.ppp(fire_week)
fire_week_ppp
```

Ensure that output is numeric.

Check for any duplicated point events.

```{r}
any(duplicated(fire_week_ppp))
```

### 6.3 Including owin object

Create a ppp object that is configured at a week level.

```{r}
week_range <- range(fire_sf$Week_num, na.rm = TRUE)
print(paste("Week range:", week_range[1], "to", week_range[2]))

st_window <- list(
  xrange = kbb_owin$xrange,  # Spatial x bounds
  yrange = kbb_owin$yrange,  # Spatial y bounds
  trange = week_range        # Temporal bounds (your Week_num range)
)

fire_stppp <- fire_week_ppp
fire_stppp$window <- st_window
```

Combine ppp and owin objects.

```{r}
fire_week_owin <- fire_week_ppp[kbb_owin]
summary(fire_week_owin)
```

### 6.4 Computing Spatio-temporal KDE

```{r}
kde_week <- spattemp.density(
  fire_week_owin)
summary(kde_week)
```

### 6.5 Plot the Spatio-temporal KDE object

KDE is plotted by week.

We can see that from Week 30 (~ end July), there are more forest fires occurring in the northern parts of the area.

From Week 38 (~ mid September), the southern parts of the area experience more forest fires, and occurrences increase until Week 46 (~ mid November), where they start to reduce in number.

```{r}
#| fig-width: 12
#| fig-height: 10
#| out.width: 100%

tims <- seq(2,51)
par(mfcol=c(2,3))
for(i in tims){ 
  plot(kde_week, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at week",i))
}
```