---
title: "2: Hot Spot & Cold Spot Analysis of Bus Ridership"
author: "Emma Chew En Chin"
date: "Oct 14, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

In this study, we utilise the following methods of **Local Measures of Spatial Autocorrelation (LMSA)** and **Emerging Hot Spot Analysis (EHSA)**, which are geospatial methodologies to analyse spatio-temporal patterns of bus ridership in Singapore.

We aim to move beyond basic mapping to identify **hotspots** (areas of high ridership concentration) and **coldspots** (areas of low ridership) . Understanding these localised patterns will provide more insight into human mobility, and infrastructure improvements for public transport planning in Singapore.

**Local Measures of Spatial Autocorrelation (LMSA)** is used to identify specific locations where values are significantly grouped. Instead of providing a single summary for the entire study area, methods such as **Local Moran's I**, **Local Geary’s C**, and **Getis-Ord Gi\***, calculate a distinct statistic for every point. From this, the local neighbourhood around the point can be tested to see if similar (or dissimiliar) values are clustered together more than random chance. The goal is to precisely map concentrated high values (hotspots) and concentrated low values (coldspots), and unusual deviations or outliers.

**Emerging Hot Spot Analysis (EHSA)** is used to identify spatio-temporal changes in spatial clustering. It takes static data and converts it into a dynamic **space-time cube** to reveal how patterns are shifting. EHSA first utilizes the local Gi∗ statistic to find where the clusters (hot and cold spots) exist, then integrates the **Mann-Kendall trend test** to determine their developmental path. This combination allows for a categorisation of trends, identifying clusters that are, for example, becoming "new," "intensifying," or "diminishing." Ultimately, EHSA provides a temporal narrative, moving beyond static analysis to explain the **dynamics and persistence** of high and low value concentrations.

### 1.1 Datasets

**Aspatial Data**

-   *Passenger Volume by Origin Destination Bus Stops* from [LTA DataMall](https://datamall.lta.gov.sg/content/datamall/en.html)

**Geospatial Data**

-   *Bus Stop Location* from LTA DataMall. It provides information about all the bus stops currently being serviced by buses, including the bus stop code (identifier) and location coordinates.

-   *Master Plan 2019 Subzone Boundary (No Sea)* from [Singapore’s open data portal](https://data.gov.sg/).

### 1.2 Methodology

**Step 1: Data Aggregation and Spatial Encoding** First, the raw bus ridership data (collected via smart cards and GPS) will be preprocessed and aggregated into a suitable spatial framework. This involves calculating the total number of passenger trips by origin and destination for defined time intervals (e.g., morning and evening peak hours, off-peak) and aggregating these counts to a defined spatial unit.

**Step 2: Local Measures of Spatial Autocorrelation (LMSA)**

Next, Local Measures of Spatial Autocorrelation (LMSA), specifically the Local Moran’s I, Local Geary’s C, and the Getis-Ord Gi\* statistics, will be used to identify statistically significant static hotspots (high-ridership clusters) and coldspots (low-ridership clusters).

**Step 3: Spatio-Temporal Analysis (EHSA)**

Emerging Hot Spot Analysis (EHSA) will be conducted by organising the static Gi∗ results from all time intervals into a space-time cube. The output will classify areas into categories such as "new," "intensifying," "diminishing," or "sporadic" hot and cold spots, providing dynamic insights into how bus ridership patterns are changing over time.

## 2 Data Preparation

### 2.1 Loading the R packages

The following R packages are used:

-   `spdep`: compute spatial weights and global and local spatial autocorrelation statistics
-   `Kendall`: conduct non-parametric trend tests and detects temporal trends in spatial patterns

```{r}
pacman::p_load(tidyverse, sf, spdep, sfdep, tmap, knitr, httr, rvest, plotly, Kendall)
```

### 2.2 Importing Data

#### 2.2.1 Bus Stop Location

As Bus Stop location data has crs = 9001, we standardise it into crs = 3414.

```{r}
BusStop = st_read(dsn = "data/LTADataMall/",
                  layer = "BusStop") %>%
  st_transform(crs = 3414)
```

#### 2.2.2 Master Plan 2019 Subzone

```{r}
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>%
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
mpsz_sf <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS", 
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")
```

```{r}
write_rds(mpsz_cl, 
          "data/rds/mpsz_cl.rds")
```

#### 2.2.3 O-D Bus Trips

```{r}
odbus <- read_csv("data/LTADataMall/origin_destination_bus_202508.csv")
```

##### 2.2.3.1 Visualise the Bus Stops

We see some bus stops outside the polygons. (Due to the JB border)

```{r}
tm_shape(mpsz_cl) + 
  tm_polygons() + 
  tm_shape(BusStop) + 
  tm_dots()
```

##### 2.2.3.2 Extract Bus Stops within Singapore

We do not use left join as we only want to get bus stops within mpsz.

```{r}
BusStop_in_SG <- st_join(
  BusStop, mpsz_cl,
  join = st_within,
  left = FALSE)
```

Plot the bus stop data again to check that the bus stops have been removed.

```{r}
tm_shape(mpsz_cl) + 
  tm_polygons() + 
  tm_shape(BusStop_in_SG) + 
  tm_dots()
```

##### 2.2.3.3 Deriving Analytical Hexagon

```{r}
hexagon <- st_make_grid(mpsz_cl,
                        cellsize = 750,
                        what = "polygon",
                        square = FALSE) %>%
  st_sf()
```

##### 2.2.3.4 Filtering hexagons with bus stops

`n_collisions` counts the number of bus stops in each hexagon. Only retain hexagons with bus stops.

```{r}
hexagon$n_collisions =
  lengths(st_intersects(
    hexagon, BusStop_in_SG))

hexagon <- filter(hexagon, n_collisions > 0)
```

##### 2.2.3.5 Assigning ids to each hexagon

Create id for each hexagon as previous hexagon table has no id.

```{r}
hexagon <- hexagon %>%
  select(,-n_collisions)

hexagon$HEX_ID <- sprintf(
  "H%04d", seq_len(nrow(hexagon))) %>%
  as.factor()

kable(head(hexagon))
```

#### 2.2.4 Preparing Trip Generation Data

We are only interested in ORIGIN_PT_CODE (trip origin).

##### 2.2.4.1 Cleaning the data

Unnecessary columns are removed

```{r}
trips <- odbus %>%
  select(c(ORIGIN_PT_CODE, DAY_TYPE, TIME_PER_HOUR, TOTAL_TRIPS)) %>%
  rename(BUS_STOP_N = ORIGIN_PT_CODE) %>%
  rename(HOUR_OF_DAY = TIME_PER_HOUR) %>%
  rename(TRIPS = TOTAL_TRIPS)
trips$BUS_STOP_N <- as.factor(trips$BUS_STOP_N)
kable(head(trips))
```

##### 2.2.4.2 Combining bus stop and hexagon layers

```{r}
bs_hex <-st_intersection(
  BusStop, hexagon) %>%
  st_drop_geometry() %>%
  select(c(BUS_STOP_N, HEX_ID))
kable(head(bs_hex))
```

##### 2.2.4.3 Adding Hex_ID into bus trips data

```{r}
trips <- inner_join(trips, bs_hex)
kable(head(trips))
```

#### 2.2.5 Create Total Trips per Hexagon dataset

This section prepares the data for **Local Measures of Spatial Association (LMSA)** analysis. We join the hexagon spatial data with the aggregated total trips data within each hexagon using **HEX_ID**. This dataset allows us to effectively visualise the spatial distribution of total bus ridership.

##### 2.2.5.1 Aggregate Trip Data by Hexagon

Within *trips* dataset, we calculate the total TRIPS per hexagon ("HEX_ID").

```{r}
trips_agg <- trips %>%
  group_by(HEX_ID) %>%
  summarise(total_trips = sum(TRIPS, na.rm = TRUE),
            .groups = 'drop')

kable(head(trips_agg))
```

##### 2.2.5.2 Relational Join

Join *trips_agg* with *hexagon* on "HEX_ID".

```{r}
hexagon_trips <- hexagon %>%
  inner_join(trips_agg, by = "HEX_ID") %>%
  mutate(total_trips = replace_na(total_trips, 0))

kable(head(hexagon_trips))
```

##### 2.2.5.3 Removing Disjointed Subgraphs from Total Trips Data

After creating the main analytical dataset, we identify and remove disjointed subgraphs to ensure a connected spatial network for accurate spatial autocorrelation analysis.

```{r}
# Create initial weights to identify connected components
hexagon_trips_weights <- hexagon_trips %>%
  mutate(nb = st_contiguity(geometry))

# Identify connected components
components <- n.comp.nb(hexagon_trips_weights$nb)
component_sizes <- table(components$comp.id)

print("Connected component sizes:")
print(component_sizes)

# Keep only the largest connected component
main_component_id <- which.max(component_sizes)
hexagon_trips <- hexagon_trips[components$comp.id == main_component_id, ]

cat("Retained component", main_component_id, "with", nrow(hexagon_trips), "hexagons\n")
cat("Removed", sum(component_sizes) - nrow(hexagon_trips), "hexagons from disjoint subgraphs\n")
```

#### 2.2.6 Create Total Trips per Hexagon and Peak Hour Period dataset

This section prepares the data for **Emerging Hot Spot Analysis (EHSA)**. We perform a spatial join to link the `hexagon` grid geometries with the aggregated bus trips data. By preserving the `HOUR_OF_DAY` dimension and calculating the total trips in each Peak Hour Period, we successfully build the spatio-temporal dataset required for constructing the space-time cube.

##### 2.2.6.1 Aggregate Trip Data by Hexagon

Within *trips* dataset, we calculate the total TRIPS per hexagon ("HEX_ID").

```{r}
trips_peak <- trips %>%
  mutate(
    # Ensure DAY_TYPE is a character/factor for matching
    DAY_TYPE = as.character(DAY_TYPE),
    # Ensure HOUR_OF_DAY is numeric for range checking (e.g., 6, 9, 17)
    HOUR_OF_DAY = as.numeric(HOUR_OF_DAY)
  ) %>%
  
  mutate(
    Peak_Period = case_when(
      # WEEKDAY Morning Peak: 6am to 9am (inclusive)
      DAY_TYPE == "WEEKDAY" & HOUR_OF_DAY >= 6 & HOUR_OF_DAY <= 9 ~ "Weekday Morning Peak",
      # WEEKDAY Afternoon Peak: 5pm to 8pm (inclusive)
      DAY_TYPE == "WEEKDAY" & HOUR_OF_DAY >= 17 & HOUR_OF_DAY <= 20 ~ "Weekday Afternoon Peak",
      # WEEKEND/HOLIDAY Peak: 11am to 8pm (inclusive)
      (DAY_TYPE == "WEEKENDS/HOLIDAY" | DAY_TYPE == "HOLIDAY") & HOUR_OF_DAY >= 11 & HOUR_OF_DAY <= 20 ~ "Weekend/Holiday Peak",
      # Everything else is Off-Peak
      TRUE ~ "Off-Peak"
    )
  )

kable(head(trips_peak))
```

```{r}
trips_agg_peak <- trips_peak %>%
  group_by(HEX_ID, Peak_Period) %>%
  summarise(
    total_trips = sum(TRIPS, na.rm = TRUE),
    .groups = 'drop'
  )

kable(head(trips_agg_peak))
```

##### 2.2.6.2 Relational Join

Join *trips_agg_peak* with *hexagon* on "HEX_ID".

```{r}
hexagon_trips_peak <- hexagon %>%
  inner_join(trips_agg_peak, by = "HEX_ID") %>%
  mutate(total_trips = replace_na(total_trips, 0))

kable(head(hexagon_trips_peak))
```

##### 2.2.6.3 Removing Disjointed Subgraphs from Peak Hour Data

We apply the same cleaning to the peak hour dataset to maintain consistency across all analyses.

```{r}
# Create initial weights to identify connected components
hexagon_trips_peak_weights <- hexagon_trips_peak %>%
  mutate(nb = st_contiguity(geometry))

# Identify connected components
components_peak <- n.comp.nb(hexagon_trips_peak_weights$nb)
component_sizes_peak <- table(components_peak$comp.id)

print("Connected component sizes for peak data:")
print(component_sizes_peak)

# Keep only the largest connected component
main_component_id_peak <- which.max(component_sizes_peak)
hexagon_trips_peak <- hexagon_trips_peak[components_peak$comp.id == main_component_id_peak, ]

cat("Retained component", main_component_id_peak, "with", nrow(hexagon_trips_peak), "hexagons\n")
cat("Removed", sum(component_sizes_peak) - nrow(hexagon_trips_peak), "hexagons from disjoint subgraphs\n")
```

#### 2.2.7 Create Total Trips per Hexagon per Hour Dataset

This section prepares the data for **Emerging Hot Spot Analysis (EHSA)**. We aggregate trips by both hexagon and hour to create a spatio-temporal dataset that will form the space-time cube.

##### 2.2.7.1 Aggregate Trip Data by Hexagon and Hour

Within *trips* dataset, we calculate the total TRIPS per hexagon ("HEX_ID") per hour ("HOUR_OF_DAY").

```{r}
trips_agg_hourly <- trips %>%
  group_by(HEX_ID, HOUR_OF_DAY) %>%
  summarise(total_trips = sum(TRIPS, na.rm = TRUE),
            .groups = 'drop')

kable(head(trips_agg_hourly))
```

##### 2.2.7.2 Relational Join

Join *trips_agg_hourly* with hexagon on "HEX_ID".

```{r}
hexagon_trips_hourly <- hexagon %>%
  inner_join(trips_agg_hourly, by = "HEX_ID") %>%
  mutate(total_trips = replace_na(total_trips, 0))

kable(head(hexagon_trips_hourly))
```

##### 2.2.7.3 Removing Disjointed Subgraphs from Hourly Data

```{r}
hexagon_trips_hourly_clean <- hexagon_trips_hourly[components$comp.id == main_component_id, ]

cat("Original hourly dataset:", nrow(hexagon_trips_hourly), "hexagon-hour combinations\n")
cat("Cleaned hourly dataset: ", nrow(hexagon_trips_hourly_clean), "hexagon-hour combinations\n")
cat("Removed", nrow(hexagon_trips_hourly) - nrow(hexagon_trips_hourly_clean), "hexagon-hour combinations from disjoint subgraphs\n")
```

#### 2.2.8 Visualising Bus Ridership in Singapore

Using **tmap**, we can plot and visualise multiple choropleth maps using equal, quantile, kmeans, headtails classification methods.

##### 2.2.8.1 Plotting Overall Bus Ridership

```{r}
small_legend_layout <- tm_layout(
  legend.text.size = 0.4, 
  legend.title.size = 0.6,
  legend.position = c("left", "bottom") 
)

equal <- tm_shape(hexagon_trips) +
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Total Bus Trips",
                position = tm_pos_in(
                  "left", "bottom"))) + 
  tm_borders(fill_alpha = 0.5) +
  tm_title("Equal interval classification") +
  small_legend_layout

quantile <- tm_shape(hexagon_trips) +
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Total Bus Trips",
                position = tm_pos_in(
                  "left", "bottom"))) + 
  tm_borders(fill_alpha = 0.5) +
  tm_title("Quantile interval classification") +
  small_legend_layout

kmeans <- tm_shape(hexagon_trips) +
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(
                style = "kmeans",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Total Bus Trips",
                position = tm_pos_in(
                  "left", "bottom"))) + 
  tm_borders(fill_alpha = 0.5) +
  tm_title("K-means interval classification") +
  small_legend_layout

headtails <- tm_shape(hexagon_trips) +
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(
                style = "headtails",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Total Bus Trips",
                position = tm_pos_in(
                  "left", "bottom"))) + 
  tm_borders(fill_alpha = 0.5) +
  tm_title("Headtails interval classification") +
  small_legend_layout

tmap_arrange(equal, 
             quantile,
             kmeans,
             headtails,
             asp=2, 
             ncol=2)
```

The visualisations show that the bus ridership data is heavily skewed, with a few areas exhibiting extremely high trip counts (a 'heavy tail' distribution).

Across the four methods, **Headtails interval classification** provides the most meaningful representation, as it effectively separates the high-ridership hot spots from the majority of lower-activity areas.

The **Headtails interval classification** map pinpoints major hotspots in the North, East, Central and West.

##### 2.2.8.2 Plotting Bus Ridership by Peak Period

We plot the Bus Ridership in Singapore by Peak Period using **Headtails interval classification** to cater to our skewed data distribution.

```{r}
# Filter data for each peak period
morning_peak <- hexagon_trips_peak %>% 
  filter(Peak_Period == "Weekday Morning Peak")

afternoon_peak <- hexagon_trips_peak %>% 
  filter(Peak_Period == "Weekday Afternoon Peak")

weekend_peak <- hexagon_trips_peak %>% 
  filter(Peak_Period == "Weekend/Holiday Peak")

off_peak <- hexagon_trips_peak %>% 
  filter(Peak_Period == "Off-Peak")

# Create headtails maps for each period
morning_map <- tm_shape(morning_peak) +
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(
                style = "headtails",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(title = "Total Trips")) + 
  tm_borders(col = "grey", alpha = 0.3) +
  tm_title("Weekday Morning Peak (6-9am)")

afternoon_map <- tm_shape(afternoon_peak) +
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(
                style = "headtails",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(title = "Total Trips")) + 
  tm_borders(col = "grey", alpha = 0.3) +
  tm_title("Weekday Afternoon Peak (5-8pm)")

weekend_map <- tm_shape(weekend_peak) +
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(
                style = "headtails",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(title = "Total Trips")) + 
  tm_borders(col = "grey", alpha = 0.3) +
  tm_title("Weekend/Holiday Peak (11am-8pm)")

off_peak_map <- tm_shape(off_peak) +
  tm_polygons(fill = "total_trips",
              fill.scale = tm_scale_intervals(
                style = "headtails",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(title = "Total Trips")) + 
  tm_borders(col = "grey", alpha = 0.3) +
  tm_title("Off-Peak Hours")

tmap_arrange(morning_map, afternoon_map, 
             weekend_map, off_peak_map,
             asp = 1, 
             ncol = 2)
```

For the Weekday Morning and Afternoon peak periods, we see a hot spot in the West and North. Whereas in the Weekend/Holiday peak and Off-peak hours, we see additional hotspots appearing in the North, and also East.

## 3 Local Measures of Spatial Autocorrelation (LMSA)

In this section, we will use **Local Measures of Spatial Autocorrelation (LMSA)**, specifically the Local Moran’s I, Local Geary’s C, and the Getis-Ord Gi\* statistics, to identify statistically significant static hotspots (high-ridership clusters) and coldspots (low-ridership clusters) for each defined time interval.

### 3.1 Computing Contiguity Spatial Weights

First, we will construct spatial weights of the study area for local spatial autocorrelation statistics. The spatial weights is used to define the neighbourhood relationships between the geographical units (i.e. hexagons) in the study area.

```{r}
wm_q <- hexagon_trips %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb, style = "W"),
         .before = 1)
```

```{r}
summary(wm_q$nb)
```

The summary report above shows that there are 799 hexagons with an average of 4.8 neighbours per hexagon. The connectivity distribution shows 352 hexagons have the maximum of 6 neighbours, while only 8 are connected with 1 neighbour.

### 3.2 Row-standardised weights matrix

Next, we will assign weights to the neighbouring polygons defined by contiguity using the spatial weights matrix (**W**).

In **Row-standardised weights**, we assign an equal fraction of influence to each neighbor, so that the sum of all weights for any given hexagon equals 1 ($1/(\text{Number of Neighbors})$).

This prevents hexagons with many neighbors from unfairly dominating the spatial autocorrelation results and is useful in our highly skewed dataset.

### 3.3 Computing Local Moran's I

To identify localised spatial patterns of similarity (clustering), we compute Local Moran's I, *Ii*, the Local Indicator of Spatial Association (LISA) statistic. This is done using the *st_local_moran()* function from the **sfdep** package.This function requires the hexagon attribute data (e.g., *total_trips*) and the row-standardised spatial weights list (*Wt*).

The resulting *Ii* value measures the degree of similarity between a hexagon and its neighbors. Its interpretation is direct:

-   A value of *Ii* \> 0 indicates strong positive autocorrelation (High-High or Low-Low clustering/similarity).
-   A value of *Ii* \< 0 indicates strong negative autocorrelation (High-Low or Low-High dissimilarity/outliers).
-   A value near 0 suggests no spatial autocorrelation.

We will use the associated p-values to generate a final LISA Cluster Map, which visually classifies all statistically significant hot spots, cold spots, and spatial outliers.

```{r}
lisa <- wm_q %>% 
  mutate(local_moran = local_moran(
    total_trips, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)

glimpse(lisa)
```

#### 3.3.1 Mapping the local Moran’s I values

```{r}
tm_shape(lisa) +
  tm_polygons(fill = "ii",
              fill.scale = tm_scale_intervals(
                style = "pretty",
                n = 5,
                values = "brewer.RdBu"),
              fill.legend = tm_legend(
                title = "Local Morans'I",
                position = tm_pos_out())) + 
  tm_borders(fill_alpha = 0.5) +
  tm_title("Local Moran's I of Bus Ridership (Queen's method)")
```

We can observe that the map is predominantly light-colored (light blue, white, and light orange), with the majority of hexagons falling into the -1 to 1 range.

There are some areas of spatial clustering (dark blue), which indicate High-High clusters, in the East, Central and West areas.

We also observe in the North area some negative spatial autocorrelation (dark orange) which may indicate High-Low clusters.

#### 3.3.2 Mapping local Moran’s I p-values

The previous map shows that there are both positive and negative Ii values. However we can plot the p-values to understand the significance of these observations.

```{r}
tm_shape(lisa) +
  tm_polygons(fill = "p_ii", 
              fill.scale = tm_scale_intervals(
                breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
                values = "-brewer.Reds"),
              fill.legend = tm_legend(
                title = "p-value",
      position = tm_pos_out())) + 
  tm_borders(fill_alpha = 0.5) +
  tm_title("p-values of Loal Moran's I of Bus Ridership (Queen's method)")
```

From the p-values in the map, we can observe that the areas with high p-value correspond to the areas of spatial clustering.

#### 3.3.3 Mapping both local Moran’s I values and p-values

```{r}
#| warning: false

ii.map <- tm_shape(lisa) +
  tm_polygons(fill = "ii",
              fill.scale = tm_scale_intervals(
                style = "pretty",
                n = 5,
                values = "brewer.RdBu"),
              fill.legend = tm_legend(
                title = "Local Moran's I",
                position = tm_pos_in(
                  "left", "bottom"))) + 
  tm_borders(fill_alpha = 0.5) +
  tm_title("Local Moran's I of Bus Ridership (Queen's method)")

p_ii.map <- tm_shape(lisa) +
  tm_polygons(fill = "p_ii", 
              fill.scale = tm_scale_intervals(
                breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
                values = "-brewer.Reds"),
              fill.legend = tm_legend(
                title = "p-value",
      position = tm_pos_in("left", "bottom")
    )) + 
  tm_borders(fill_alpha = 0.5) +
  tm_title("p-values of Local Moran's I of Bus Ridership (Queen's method)")

tmap_arrange(ii.map, p_ii.map, asp=1, ncol=2)
```

#### 3.3.4 Preparing and Visualising LISA Map

LISA maps includes choropleth maps showing LISA values, Moran scatterplots which depict local spatial autocorrelation for individual locations, and hot spot analysis maps.

##### 3.3.4.1 Plotting Moran scatterplot

The Moran scatterplot visualises the relationship between a variable’s values and the average values of its neighboring locations (its spatial lag) to identify spatial autocorrelation.

`st_lag()` of **sfdep** package is used to calculates the spatial lag of GDPPC given a neighbor (i.e. nb) and weights list (i.e. wt).

```{r}
lisa <- lisa %>%
  mutate(lag_trips = st_lag(
    total_trips, nb, wt),
    .before = 1) %>%
  unnest(lag_trips)
```

The Moran Scatterplot of total_trips is plotted against its spatial lag version.

```{r}
#| warning: false

ggplot(data = lisa, 
       aes(x = total_trips, 
           y = lag_trips)) +
  geom_point() +
  geom_smooth(method = "lm", 
              se = FALSE, 
              color = "red") +
  labs(x = "Total_Trips",
       y = "Spatial Lag of Total_Trips",
       title = "Moran Scatterplot") +
  theme_minimal()
```

From the scatterplot, we can observe that the trend line shows a positive relationship between Total_Trips and its spatial lag version, indicating positive global spatial autocorrelation.

```{r}
ggplot(data = lisa, 
       aes(x = total_trips, 
           y = lag_trips, 
           color = mean)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              color = "black") +
  geom_hline(yintercept=mean(lisa$lag_trips), lty=2) + 
    geom_vline(xintercept=mean(lisa$total_trips), lty=2) +
  scale_color_manual(
    values = c("High-High" = "red", 
               "Low-Low" = "blue",
               "Low-High" = "lightblue", 
               "High-Low" = "pink")) +
  labs(x = "Total_Trips",
       y = "Spatial Lag of Total_Trips",
       title = "Moran Scatterplot with LISA Quadrants") +
  theme_minimal()

```

From the Moran scatterplot with LISA Quadrants, we can understand that Low-Low and High-High clusters are more prevalant, and there is a high incidence of prominent outliers in the form of Low-High clusters.

##### 3.3.4.2 Plotting Moran scatterplot with standardised variable

scale() is used to center and scale the variable. Centering is implemented by subtracting the mean (omitting NAs), and scaling is done by dividing the (centered) variable by their standard deviations.

```{r}
lisa <- lisa %>%
  mutate(z_trips = scale(total_trips),
         z_lag_trips = scale(lag_trips),
         .before = 1)
```

```{r}
#| warning: false

ggplot(data = lisa, 
       aes(x = z_trips, 
           y = z_lag_trips, 
           color = mean)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              color = "black") +
  geom_hline(yintercept=mean(lisa$z_lag_trips), lty=2) + 
    geom_vline(xintercept=mean(lisa$z_trips), lty=2) +
  scale_color_manual(
    values = c("High-High" = "red", 
               "Low-Low" = "blue",
               "Low-High" = "lightblue", 
               "High-Low" = "pink")) +
  labs(x = "Standardised Total_Trips",
       y = "Standardised Spatial Lag of Total_Trips",
       title = "Standardised Moran Scatterplot with LISA Quadrants") +
  theme_minimal()
```

After applying standardisation, the output findings remain the same.

##### 3.3.4.3 Preparing LISA map classes

We will set a statistical significance level for the local Moran.

```{r}
signif <- 0.05  
```

##### 3.3.4.4 Plotting and visualising LISA map

By default, the LISA classes are stored in *mean*, *median* and *pysal* fields of lisa sf data frame. To prepare the LISA map, we will derive a new field called LISA-cluster from either one of these three fields.

A new class 'insignificant' has been added for records where p_ii \> 0.05.

```{r}
lisa <- lisa %>%
  mutate(
    LISA_cluster = ifelse(
      p_ii < 0.05,
      as.character(mean), 
      "Insignificant"),
    LISA_cluster = factor(
      LISA_cluster,
      levels = c("Insignificant", "Low-Low", "Low-High", "High-Low", "High-High")
    )
  )
```

```{r}
#| warning: false

lisa <- lisa %>%
  mutate(
    moran_categorized = case_when(
      p_ii > 0.05 ~ "Not Significant",
      ii > 2 ~ "Strong Positive",
      ii > 1 & ii <= 2 ~ "Moderate Positive",
      ii > -0.5 & ii <= 1 ~ "Weak/Random", 
      ii > -1 & ii <= -0.5 ~ "Moderate Negative",
      ii <= -1 ~ "Strong Negative",
      TRUE ~ "Not Significant"
    ),
    moran_categorized = factor(
      moran_categorized,
      levels = c(
        "Strong Positive",     
        "Moderate Positive",   
        "Weak/Random",         
        "Moderate Negative",   
        "Strong Negative",     
        "Not Significant"      
      )
    )
  )


ii.map <- tm_shape(lisa) +
  tm_polygons(
    fill = "moran_categorized",
    fill.scale = tm_scale_categorical(
      values = c(
        "Strong Positive" = "#b2182b",    # Dark red
        "Moderate Positive" = "#ef8a62",  # Medium red
        "Weak/Random" = "#fddbc7",        # Light red/beige
        "Moderate Negative" = "#67a9cf",  # Medium blue
        "Strong Negative" = "#2166ac",    # Dark blue
        "Not Significant" = "grey80"      # Grey
      )
    ),
    fill.legend = tm_legend(
      title = "Local Moran's I\n(p < 0.05)",
      position = tm_pos_in("left", "bottom")
    )
  ) + 
  tm_borders(fill_alpha = 0.5) +
  tm_title("Local Moran's I Values (p < 0.05)")

lisa_map <- tm_shape(lisa) + 
  tm_polygons(
    fill = "LISA_cluster",
    fill.scale = tm_scale_categorical(
      values = c(
        "grey80",      # Insignificant
        "blue",        # Low-Low
        "lightblue",   # Low-High
        "pink",        # High-Low
        "red"          # High-High
      )
    ),
    fill.legend = tm_legend(
      title = "LISA Cluster",
      position = tm_pos_in("left", "bottom"))
  ) +
  tm_borders() +
  tm_title("Local Moran's I Clusters (p < 0.05)")

tmap_arrange(ii.map, lisa_map, asp=1, ncol=2)
```

From local Moran's I values, we can see strong positive autocorrelation in the West, Northeast and East.

We can observe many statistically significant high-high clusters across the map appearing distinctly in the East/Northeast, Central, and isolated pockets in the West.

### 3.4 Computing Local Geary's C

To assess the presence of localised spatial patterns of dissimilarity, we compute Local Geary's C, *Ci*, using the `st_local_geary()` function from the **sfdep** package.

The resulting *Ci* value provides a measure of dissimilarity between a hexagon and its neighbors. Its interpretation is inverse to Local Moran's I:

-   A value of *Ci* close to 0 indicates strong positive autocorrelation (High-High or Low-Low clustering/similarity).
-   A value of *Ci* close to 2 indicates strong negative autocorrelation (High-Low or Low-High dissimilarity/outliers).
-   A value near 1 suggests no spatial autocorrelation.

#### 3.4.1 Creating Local Geary's C statistics

First, we apply log transformation to *total_trips* as it is right skewed.

A right-skewed distribution will affect Geary's C calculation, as very high values will dominate, affecting the squared difference calculation, and local clustering of moderate values may not be detected.

```{r}
# Create the transformed data if it doesn't exist
if (!exists("wm_q_transformed")) {
  wm_q_transformed <- wm_q %>%
    mutate(log_trips = log(total_trips + 1))
}

# Use only wm_q and wm_q_transformed for the comparison
combined_data <- bind_rows(
  wm_q %>% 
    select(HEX_ID, total_trips, geometry) %>%
    mutate(type = "Original", value = total_trips),
  wm_q_transformed %>% 
    select(HEX_ID, total_trips, log_trips, geometry) %>%
    mutate(type = "Log-Transformed", value = log_trips)
)

ggplot(combined_data, aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(~type, scales = "free") +
  labs(title = "Distribution Comparison of Total_Trips", x = "Value")
```

```{r}
local_geary_results <- wm_q_transformed %>%
  mutate(
    local_geary = local_c_perm(
      x = log_trips,  # Use the transformed variable
      nb = nb,
      wt = wt, 
      nsim = 999 
    )
  )

local_geary_data <- local_geary_results %>%
  select(HEX_ID, total_trips, log_trips, local_geary) %>%  
  tidyr::unnest(local_geary)

glimpse(local_geary_data)
```

```{r}
local_geary_data <- local_geary_data %>%
  filter(HEX_ID %in% hexagon_trips$HEX_ID)

cat("local_geary_data now has:", nrow(local_geary_data), "hexagons\n")
```

#### 3.4.2 Visualising Local Geary's C Values and clusters

```{r}
# Create Geary_Cluster column
local_geary_data <- local_geary_data %>%
  mutate(
    Geary_Cluster = case_when(
      p_ci >= 0.05 ~ "Insignificant",
      ci < 1 & cluster == "High-High" ~ "High-High Similarity",
      ci < 1 & cluster == "Low-Low" ~ "Low-Low Similarity", 
      ci > 1 & cluster == "High-Low" ~ "High-Low Dissimilarity",
      ci > 1 & cluster == "Low-High" ~ "Low-High Dissimilarity",
      TRUE ~ "Insignificant"
    ),
    # Convert to factor for consistent ordering
    Geary_Cluster = factor(
      Geary_Cluster,
      levels = c(
        "High-High Similarity",
        "Low-Low Similarity",
        "High-Low Dissimilarity", 
        "Low-High Dissimilarity",
        "Insignificant"
      )
    ),
    # Create strength classification for significant values
    geary_strength = case_when(
      p_ci >= 0.05 ~ "Not Significant",
      ci < 0.5 ~ "Strong Positive Similarity",
      ci >= 0.5 & ci < 1 ~ "Weak Positive Similarity",
      ci >= 1 & ci < 1.5 ~ "Weak Negative Similarity", 
      ci >= 1.5 ~ "Strong Negative Similarity",
      TRUE ~ "Not Significant"
    ),
    geary_strength = factor(
      geary_strength,
      levels = c(
        "Strong Positive Similarity",
        "Weak Positive Similarity",
        "Weak Negative Similarity",
        "Strong Negative Similarity",
        "Not Significant"
      )
    )
  )

# Map 1: Geary's C Strength values
geary_strength_map <- tm_shape(local_geary_data) +
  tm_polygons(
    fill = "geary_strength",
    fill.scale = tm_scale_categorical(
      values = c(
        "Strong Positive Similarity" = "#2166ac",
        "Weak Positive Similarity" = "#67a9cf",
        "Weak Negative Similarity" = "#fddbc7", 
        "Strong Negative Similarity" = "#d6604d",
        "Not Significant" = "grey80"
      )
    ),
    fill.legend = tm_legend(title = "Geary's C Strength")
  ) +
  tm_borders(col = "black", fill_alpha = 0.3) +
  tm_title("Local Geary's C Strength (p < 0.05)")

# Map 2: Geary Clusters (custom classification)
geary_clusters_map <- tm_shape(local_geary_data) +
  tm_polygons(
    fill = "Geary_Cluster",
    fill.scale = tm_scale_categorical(
      values = c(
        "High-High Similarity" = "lightblue",
        "Low-Low Similarity" = "pink",  
        "High-Low Dissimilarity" = "blue",
        "Low-High Dissimilarity" = "red",
        "Insignificant" = "grey80"
      )
    ),
    fill.legend = tm_legend(title = "Geary Similarity Clusters")
  ) +
  tm_borders(col = "black", fill_alpha = 0.3) +
  tm_title("Local Geary's C Similarity Patterns")

# Arrange maps side by side
tmap_arrange(geary_strength_map, geary_clusters_map, 
             ncol = 2, 
             heights = c(1, 1))
```

From the visualisations of Local Geary's C values, we can see that the bus ridership on the West and North exhibits **strong positive similarity** (dark blue shades), where majority of hexagons have *total_trips* similar to their neighbours (i.e. Low-Low or High-High clustering). Whereas in the South and Northeast, there exists some areas of **strong negative similarity** (dark orange shades).

A closer look at the Geary clusters show that the West and North areas exhibit Low-High dissimilarity (pink shades), which means that there are cold spots in these areas.

### 3.5 Computing Getis-Ord Gi\*

To identify localised spatial clusters of extreme values (hot spots and cold spots), we compute the Getis-Ord Gi\* statistic. This is achieved using the `st_local_g()` function from the **sfdep** package. The resulting Gi\* value provides a measure of similarity (clustering), specifically comparing a hexagon's value to its neighbours, including itself.

Its interpretation is as such:

-   A positive and large Gi\* value indicates positive autocorrelation (High-High clustering/Hot Spot).
-   A negative and large Gi\* value indicates positive autocorrelation (Low-Low clustering/Cold Spot).
-   A Gi\* value near 0 indicates no spatial autocorrelation; the values are randomly distributed.

We will use the associated p-values and the Z-scores derived from Gi\* to generate a final map that visually pinpoints the statistically significant hot spots and cold spots of bus ridership.

#### 3.5.1 Computing distance weight matrix

Getis-Ord Gi\* analysis defines neighbours based on distance, and there are 2 types of distance-based proximity matrices that can be utilised:

-   fixed distance weight matrix,
-   adaptive distance weight matrix

##### 3.5.1.1 Computing adaptive distance weights

For bus ridership analysis in Singapore, fixed distance weights capture **natural variations** where dense hexagons have more neighbouring stops within a fixed radius, smoothing local relationships.

In contrast, adaptive distance weights ensure consistent analysis by maintaining the same number of neighbors (k-nearest) for each hexagon, providing **uniform spatial context** regardless of urban density, which helps isolate true ridership patterns from structural biases of the transport infrastructure.

Hence, we will use adaptive distance weights for Gi\* analysis.

```{r}
hexagon_trips_adw <- hexagon_trips %>%
  mutate(nb = include_self(
    st_knn(
      st_geometry(geometry),
      k = 6)),
    wt = st_weights(
      nb, style = "W"),
    .before = 1)
```

#### 3.5.2 Computing local Gi\*

Local Gi\* is a Getis-Ord statistic that identifies statistically significant hot spots (clusters of high values) and cold spots (clusters of low values) in spatial data. It calculates a Z-score and p-value for each location by comparing local neighborhood sums to the global average, determining whether observed spatial clustering occurs by random chance.

```{r}
HCSA_adw <- hexagon_trips_adw %>%
  mutate(
    gistar = local_gstar_perm(
      total_trips, nb, wt, nsim = 99),
    .before = 1) %>%
  unnest(gistar)
```

#### 3.5.3 Preparing and Visualising HCSA Map

##### 3.5.3.1 Mapping Gi\* with adaptive distance weights

Thematic mapping functions of tmaps are used to prepare a choropleth map showing the distribution of Gi\* values.

```{r}
# Create Gi* intervals
HCSA_adw <- HCSA_adw %>%
  mutate(
    gi_strength = case_when(
      p_value > 0.05 ~ "Not Significant",
      gi_star >= 3 ~ "Very Strong Hot Spot",
      gi_star >= 2 & gi_star < 3 ~ "Strong Hot Spot", 
      gi_star >= 1 & gi_star < 2 ~ "Moderate Hot Spot",
      gi_star > 0 & gi_star < 1 ~ "Weak Hot Spot",
      gi_star <= -3 ~ "Very Strong Cold Spot",
      gi_star <= -2 & gi_star > -3 ~ "Strong Cold Spot",
      gi_star <= -1 & gi_star > -2 ~ "Moderate Cold Spot",
      gi_star < 0 & gi_star > -1 ~ "Weak Cold Spot",
      TRUE ~ "Not Significant"
    ),
    # Convert to factor with fixed order
    gi_strength = factor(
      gi_strength,
      levels = c(
        "Very Strong Hot Spot",
        "Strong Hot Spot", 
        "Moderate Hot Spot",
        "Weak Hot Spot",
        "Weak Cold Spot",
        "Moderate Cold Spot",
        "Strong Cold Spot",
        "Very Strong Cold Spot",
        "Not Significant"
      )
    )
  )

# Map with fixed legend order
tm_shape(HCSA_adw) +
  tm_polygons(
    fill = "gi_strength",
    fill.scale = tm_scale_categorical(
      values = c(
        "Very Strong Hot Spot" = "#8B0000",
        "Strong Hot Spot" = "#FF0000", 
        "Moderate Hot Spot" = "#FF6B6B",
        "Weak Hot Spot" = "#FFB6C1",
        "Very Strong Cold Spot" = "#00008B",
        "Strong Cold Spot" = "#0000FF",
        "Moderate Cold Spot" = "#6B8EFF", 
        "Weak Cold Spot" = "#B6C1FF",
        "Not Significant" = "grey80"
      )
    ),
    fill.legend = tm_legend(title = "Gi* Cluster Strength")
  ) +
  tm_borders(col = "grey", alpha = 0.3) +
  tm_title("Gi* Hot and Cold Spots by Intensity (p ≤ 0.05)")
```

From the visualisation map, we can see that there are statistically significant **Very Strong Hot Spots** in the West, North and East areas (Gi\*\>0). The Gi\* values show that high bus ridership values in these regions is unlikely to occur by random chance (p \< 0.05), representing genuine spatial patterns rather than random distributions.

## 4 Emerging Hot Spot Analysis (EHSA)

**Emerging Hot Spot Analysis (EHSA)** is a spatio-temporal technique that identifies how bus ridership clusters are changing across Singapore's hexagon grid over time.

The process works by first converting peak ridership data into a space-time cube. For each period, the Getis-Ord Gi\* statistic identifies Hot and Cold Spots. The Mann-Kendall trend test is then applied to the sequence of these Gi\* values for every hexagon to categorise areas of new/growing/persistent hot/cold spots, providing deeper insight for bus demand planning.

### 4.1 Creating a Space-Time Cube

We use `spacetime()` of **sfdep** package to create an spatio-temporal cube, and verify its object type.

#### 4.1.1 Cleaning hexagon_trips_peak and hexagon Datasets

```{r}
# Add numeric time_id to establish temporal order
hexagon_trips_peak <- hexagon_trips_peak %>%
  mutate(
    time_id = case_when(
      Peak_Period == "Weekday Morning Peak" ~ 1L,
      Peak_Period == "Weekday Afternoon Peak" ~ 2L,
      Peak_Period == "Weekend/Holiday Peak" ~ 3L,
      Peak_Period == "Off-Peak" ~ 4L,
      TRUE ~ NA_integer_  # Handle any unexpected values
    )
  ) %>%
  # Reorder columns for better readability
  select(HEX_ID, time_id, Peak_Period, total_trips, geometry)

# Show the updated data with time_id
kable(head(hexagon_trips_peak))
```

```{r}
kable(head(hexagon))
```

```{r}
trips_peak_st <- spacetime(hexagon_trips_peak, hexagon,
                      .loc_col = "HEX_ID",
                      .time_col = "time_id")
is_spacetime_cube(trips_peak_st)
```

From the initial run of Space-Time cube, there is an error stating that some hexagons does not have data for all 4 time periods.

A perfect cube must have this property: total rows = (number of unique HEX_IDs) × (number of unique time_id values)

```{r}
n_locations <- n_distinct(hexagon_trips_peak$HEX_ID)
n_timeperiods <- n_distinct(hexagon_trips_peak$time_id)
n_rows <- nrow(hexagon_trips_peak)

cat("Locations:", n_locations, "\n")
cat("Time periods:", n_timeperiods, "\n")
cat("Rows:", n_rows, "\n")
cat("Expected rows:", n_locations * n_timeperiods, "\n")
```

From the result, we observe that the existing rows is fewer than expected rows. Hence we print missing combinations of HEX_ID and time_id.

```{r}
missing_combinations <- hexagon_trips_peak %>%
  count(HEX_ID, time_id) %>%
  complete(HEX_ID, time_id) %>%
  filter(is.na(n))

cat("Missing combinations:", nrow(missing_combinations), "\n")
print(missing_combinations)
```

We will fill in the missing hexagons with 0 values.

```{r}
# Get all possible combinations
all_combinations <- tidyr::crossing(
  HEX_ID = unique(hexagon_trips_peak$HEX_ID),
  time_id = unique(hexagon_trips_peak$time_id)
)

# Merge and fill missing with 0
hexagon_trips_peak_balanced <- all_combinations %>%
  left_join(hexagon_trips_peak, by = c("HEX_ID", "time_id")) %>%
  mutate(total_trips = ifelse(is.na(total_trips), 0, total_trips)) %>%
  arrange(HEX_ID, time_id)
```

We will then rebuild the Space-Time cube.

```{r}
trips_peak_st <- spacetime(
  hexagon_trips_peak_balanced,
  hexagon,
  .loc_col = "HEX_ID",
  .time_col = "time_id"
)

is_spacetime_cube(trips_peak_st)
```

### 4.2 Computing Gi\*

#### 4.2.1 Deriving the spatial weights

We will derive the inverse distance weights as below.

```{r}
#| warning: false

trips_peak_nb <- trips_peak_st %>%
  activate("geometry") %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wt = st_inverse_distance(nb, 
                             geometry, 
                             scale = 1,
                             alpha = 1),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")
```

#### 4.2.2 Calculating local Gi\*

To calculate local Gi\* statistics for each time period, we group the data by *time_id* and apply the `local_gstar_perm()` function, then unnest the results.

```{r}
gi_stars <- trips_peak_nb %>% 
  group_by(time_id) %>% 
  mutate(gi_star = local_gstar_perm(
    total_trips, nb, wt)) %>% 
  tidyr::unnest(gi_star)
```

### 4.3 Mann-Kendall Test

The Mann-Kendall test detects monotonic trends in bus ridership patterns over time:

-   H₀: No monotonic trend in ridership
-   H₁: Monotonic trend present

Interpretation:

-   Reject H₀ if p-value \< significance level (e.g., 0.05)
-   Tau statistic (-1 to 1):
    -   -1: Perfectly decreasing ridership trend
    -   +1: Perfectly increasing ridership trend
    -   0: No trend

This identifies whether bus ridership is systematically increasing or decreasing across the analyzed time periods.

#### 4.3.1 Mann-Kendall Test on Gi\*

We first look at 1 hexagon, H0001.

```{r}
cbg <- gi_stars %>% 
  ungroup() %>% 
  filter(HEX_ID == "H0001") |> 
  select(HEX_ID, time_id, gi_star)
```

```{r}
ggplot(data = cbg, 
       aes(x = time_id, 
           y = gi_star)) +
  geom_line() +
  theme_light()
```

#### 4.3.2 Interactive Mann-Kendall Plot

```{r}
p <- ggplot(data = cbg, 
       aes(x = time_id, 
           y = gi_star)) +
  geom_line() +
  theme_light()

ggplotly(p)
```

#### 4.3.3 Printing Mann-Kendall test report

```{r}
cbg %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)))) %>% 
  tidyr::unnest_wider(mk)
```

In the above result, sl is the **p-value**. As p-value > 0.05, we will not reject the null hypothesis that there is no monotonic trend in ridership.

#### 4.3.4 Mann-Kendall test dataframe

We will replicate the study for the entire dataframe by grouping by *HEX_ID*.

```{r}
ehsa <- gi_stars %>%
  group_by(HEX_ID) %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)))) %>%
  tidyr::unnest_wider(mk)
head(ehsa)
```

To showcase emerging hot/cold spots, we can sort by significance level and tau.

```{r}
emerging <- ehsa %>% 
  arrange(sl, abs(tau)) %>% 
  slice(1:10)
head(emerging)
```

### 4.4 Performing Emerging Hotspot Analysis

We will perform EHSA analysis by using `emerging_hotspot_analysis()` of **sfdep** package. 

The `emerging_hotspot_analysis()` function requires:
-   x: spacetime object (e.g., trips_peak_st)
-   .var: variable name (e.g., "total_trips")
-   k: time lags (default: 1)
-   nsim: number of simulations

```{r}
ehsa <- emerging_hotspot_analysis(
  x = trips_peak_st, 
  .var = "total_trips", 
  k = 1, 
  nsim = 99
)
```

#### 4.4.1 Visualising the distribution of EHSA classes

We can identify the prevalence of each EHSA class.

```{r}
ggplot(data = ehsa,
       aes(x = classification)) +
  geom_bar()
```

From the results, we can identify that for most of the hexagons, there was no pattern identified. 

#### 4.4.2 Visualising EHSA

We will join ehsa and hexagon data using HEX_ID, to visualise the geographic distribution of EHSA classes.

```{r}
hexagon_ehsa <- hexagon %>%
  left_join(ehsa,
            by = join_by(HEX_ID == location))
```

```{r}
ehsa_sig <- hexagon_ehsa %>%
  filter(p_value < 0.05) %>%
  st_as_sf()

tmap_mode("plot")

if (nrow(ehsa_sig) > 0) {
  tm_shape(hexagon_ehsa) +
    tm_polygons() +
    tm_borders(fill_alpha = 0.5) +
    tm_shape(ehsa_sig) +
    tm_fill("classification") + 
    tm_borders(fill_alpha = 0.4)
} else {
  cat("⚠️ No significant EHSA clusters (p < 0.05), plot skipped.\n")
}
```

There is an error occurring, and we conduct a check on the significant values and find that there are no polygons where p_value < 0.05.

```{r}
nrow(ehsa_sig)
```

As Peak Hour Period dataset may have not enough datapoints for EHSA (only 4 time values), we can consider looking at trips per hour instead.
