---
title: "3: Modelling HDB Resale Prices with Geographically Weighted Methods"
author: "Emma Chew En Chin"
date: "Oct 26, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

In Singapore, understanding the factors that influence the potential prices of HDB resale flats is crucial. For prospective homeowners, this knowledge is vital as it directly impacts the long-term financial sustainability of the purchase, including the amount of housing equity they can secure and how well the flat retains its value over time. This understanding allows buyers to make informed and prudent housing decisions regarding the flat size and location that best meet their family's needs and financial standing.

This study utilises **Geographically Weighted Models (GWMs)** to build a price prediction model to analyse the key factors that influence HDB resale prices, specifically for **4-room flats** due to its popularity with young families. From this, we can produce local coefficients for each data point (or small region), reflecting the spatial heterogeneity in the data. The data period for analysis will be 1 January 2025 to 30 September 2025.

## 2 Installing and Loading R packages

::: callout-note
**Data Acquisition & Handling**

-   `httr`: Makes HTTP requests to APIs; used for url call from OneMap.
-   `jsonlite`: Parses JSON data returned from web APIs into usable R data frames.
-   `rvest`: For scraping of data from web pages.
-   `tidyverse`: A meta-package for data manipulation (dplyr), visualization (ggplot2), and more. The core toolkit for data wrangling.

**Geospatial Core**

-   `sf`: Handles spatial data (points, polygons). Used for spatial joins, distance calculations, and mapping.
-   `tmap`: Creates thematic maps (chloropleth, point maps) to visualise spatial patterns in prices and model results.

**Geospatial Modelling & Analysis**

-   `spdep`: Provides spatial dependence tests (e.g., Moran's I for spatial autocorrelation) and classic spatial econometric models.
-   `GWmodel`: The primary package for GWR and multiscale GWR. Fits models where coefficients vary geographically.
-   `SpatialML`: Provides geographically weighted random forests (gwrf()), blending random forests with local modeling.
-   `spatialRF`: Provides spatial random forest models and spatial cross-validation techniques for improved spatial prediction.

**Model Pipeline and Evaluation**

-   `rsample`: Creates training/test splits for model validation using techniques like cross-validation.
-   `Metrics`: Calculates model performance metrics (RMSE, MAE, R²) to evaluate prediction accuracy.
-   `randomForestExplainer`: Provides tools for interpreting and explaining random forest models, including variable importance and interaction effects.

**Reporting & Visualisation**

-   `knitr`: Dynamic report generation and reproducible research; integrates R code with documentation.
-   `kableExtra`: Enhances table formatting and styling in R Markdown documents and Shiny applications.

**Utility**

-   `progress`: Adds progress bars to loops and long-running operations, improving user experience during spatial computations.
:::

```{r}
pacman::p_load(httr, sf, tmap, jsonlite, progress,
               spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse, rvest,
               knitr, kableExtra, spatialRF,
               randomForestExplainer)
```

## 3 Data Preparation

### 3.1 The Datasets

::: callout-note
**Aspatial dataset:**

-   HDB Resale data: a list of HDB resale transacted prices in Singapore from Jan 2017 onwards.

**Geospatial dataset:**

-   MP14_SUBZONE_WEB_PL: a polygon feature data providing information of URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format.

-   Transport Factors:

    -   *MRT* data is a list of MRT/LRT stations in Singapore with the station names and codes. It is in geojson format.
    -   *Bus stops* data is a list of bus stops in Singapore. It is in geojson format.

-   Education Factors:

    -   *Preschool* data is a list of preschools in Singapore. It is in geojson format.
    -   *Childcare service* data is a list of childcare services in Singapore. It is in geojson format.
    -   *School* data is a list of primary, secondary, junior colleges schools in Singapore. It is in csv format.

-   Elderly-care Factors:

    -   *Eldercare* data is a list of eldercare in Singapore. It is in geojson format.

-   Recreational Factors

    -   *Parks* data is a list of parks in Singapore. It is in geojson format.
    -   *Community clubs* data is a list of community clubs in Singapore. It is in geojson format.
    -   *Sports Facilities* data is a list of public sports facilities in Singapore. It is in geojson format.
    -   *GymsSG* data is a list of public gym facilities in Singapore. It is in geojson format.

-   Food Factors

    -   *Hawker Centre* data is a list of hawker centres in Singapore. It is in geojson format.
    -   *Supermarket* data is a list of supermarkets in Singapore. It is in geojson format.

-   Healthcare Factors

    -   *CHAS clinics* data is a list of CHAS clinics in Singapore. It is in geojson format.
:::

### 3.2 Importing Shapefile Polygon data

Shapefile Polygon data, mpsz, is imported for visual checks for geospatial data to ascertain the following:

-   **Spatial coverage validation:** To verify all points fall within expected boundaries
-   **Coordinate system alignment:** To ensure that both layers use the same CRS
-   **Spatial distribution patterns:** To quickly identify clusters, gaps or unusual patterns
-   **Data completeness verification:** To find missing or outlier points

```{r}
mpsz = st_read(dsn = "data/geospatial", 
               layer = "MP14_SUBZONE_WEB_PL") %>%
  st_transform(crs = 3414)
```

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

### 3.2 Importing HDB Resale data

Next, we import and clean the HDB resale dataset by using OneMap reverse geocoding to obtain the coordinates from the street addresses of the HDB flats.

We can reverse geocode the HDB postal code, as there is only blk number and street name in the Resale Flats dataset.

This is the documentation: [OneMap Reverse Geocoding](https://www.onemap.gov.sg/apidocs/reverseGeocode)

```{r}
#| eval: false

HDBresale <-read_csv("data/rawdata/HDBresale.csv") %>%
  filter(flat_type == "4 ROOM") %>%
  filter(month >= "2025-01" & month < "2025-09")

glimpse(HDBresale)
```

Changing ST to SAINT so that it does not look like 'Street'.

```{r}
#| eval: false

HDBresale$street_name <- gsub(
  "ST\\.", "SAINT",
  HDBresale$street_name)
```

```{r}
#| eval: false

geocode <- function(block, streetname) {
  base_url <- "https://onemap.gov.sg/api/common/elastic/search"
  address <- paste(block, streetname, sep = " ")
  query <- list("searchVal" = address,
                "returnGeom" = "Y",
                "getAddrDetails" = "N",
                "pageNum" = "1")
  
  res <-GET(base_url, query = query)
  restext <- content(res, as="text")
  
  output <- fromJSON(restext) %>%
    as.data.frame %>%
    select(results.LATITUDE, results.LONGITUDE)
  
  return(output)
}
```

```{r}
#| eval: false
#| warning: false
#| results: 'hide'
#| message: false

HDBresale$LATITUDE <- 0
HDBresale$LONGITUDE <- 0

for (i in 1:nrow(HDBresale)){
  temp_output <- geocode(HDBresale[i, 4], HDBresale[i, 5])
  
  HDBresale$LATITUDE[i] <- temp_output$results.LATITUDE
  HDBresale$LONGITUDE[i] <- temp_output$results.LONGITUDE
}
```

```{r}
#| eval: false

glimpse(HDBresale)
```

Since HDBresale contains LATITUDE and LONGITUDE columns (which are in the WGS 84 geographic coordinate system, or EPSG:4326), we need to convert it into an sf (Simple Features) object and then transform it to the SVY21 (Singapore Projected Coordinate System, EPSG:3414).

```{r}
#| eval: false

HDBresale_sf <- HDBresale %>%
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs=4326) %>%
  st_transform(crs = 3414)

saveRDS(HDBresale_sf, file = "data/rds/HDBresale_sf.rds")
```

```{r}
HDBresale_sf <- read_rds("data/rds/HDBresale_sf.rds")
```

We check if there are overlapping points in HDBresale.

```{r}
overlapping_points <- HDBresale_sf %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
```

Using `jitter()` to move point features away from each other by 1m to avoid overlapping points.

```{r}
HDBresale_sf <- HDBresale_sf %>%
  st_jitter(amount = 1)
```

```{r}
plot(st_geometry(mpsz), col = NA, border = "gray", main = "HDB Resale Locations")
plot(st_geometry(HDBresale_sf), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

### 3.3 Importing Schools data

As Schools data is in address format, we also conduct the same reverse geocoding with OneMap API to obtain the coordinate data.

```{r}
schools <-read_csv("data/rawdata/Schools.csv")
```

We utilise the *postal_code* attribute in *schools* dataframe to retrieve the coordinates from OneMap API.

```{r}
#| eval: false

schools_postcodes <- unique(schools$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

# Found: geocoded, Not found: cannot be geocoded (e.g.) New development area or inaccurate postal code
found <- data.frame()
not_found <- data.frame(postcode = character(), stringsAsFactors = FALSE)

for (pc in schools_postcodes) {
  # Convert postal code to character for the API query
  query <- list(
    searchVal = as.character(pc),  # Convert to character here
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    results <- bind_rows(lapply(json$results, as.data.frame))
    results$input_postcode <- as.character(pc)  # Convert to character here too
    found <- bind_rows(found, results)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = as.character(pc), stringsAsFactors = FALSE))
  }
  Sys.sleep(0.1)
}
```

```{r}
#| eval: false

schools_clean <- schools %>%
  filter(!postal_code %in% not_found$postcode)

cat("Original schools rows:", nrow(schools), "\n")
cat("Cleaned schools rows:", nrow(schools_clean), "\n")
cat("Number of postal codes removed:", nrow(schools) - nrow(schools_clean), "\n")
```

```{r}
#| eval: false

schools_joined <- schools_clean %>%
  mutate(postal_code = as.character(postal_code)) %>%
  left_join(
    found %>%
      group_by(input_postcode) %>%
      slice(1) %>%  # Take the first geocoded result for each postal code
      ungroup(),
    by = c('postal_code' = 'input_postcode')
  ) %>%
  mutate(
    LONGITUDE = as.numeric(LONGITUDE),  
    LATITUDE = as.numeric(LATITUDE)
  ) %>%
  group_by(LONGITUDE, LATITUDE) %>%
  mutate(
    n_coords = n(),
    jitter_applied = n_coords > 1 & !is.na(LONGITUDE) & !is.na(LATITUDE),
    LONGITUDE = ifelse(jitter_applied, jitter(LONGITUDE), LONGITUDE),
    LATITUDE = ifelse(jitter_applied, jitter(LATITUDE), LATITUDE)
  ) %>%
  ungroup() %>%
  select(-n_coords)

saveRDS(schools_joined, file = "data/rds/schools_joined.rds")
```

```{r}
schools_joined <- read_rds("data/rds/schools_joined.rds")

schools_sf <- schools_joined %>%
  filter(!is.na(LONGITUDE) & !is.na(LATITUDE)) %>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

# Transform to same CRS as mpsz
schools_sf <- st_transform(schools_sf, st_crs(mpsz))
```

```{r}
plot(st_geometry(mpsz), col = NA, border = "gray", main = "School Locations")
plot(st_geometry(schools_sf), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

### 3.4 Importing MRT data

```{r}
#| eval: false

mrt <- st_read("data/geospatial/LTAMRTStationExitKML.kml") %>%
  st_transform(crs = 3414)

saveRDS(mrt, file = "data/rds/mrt_data.rds")
```

```{r}
mrt_sf <- read_rds("data/rds/mrt_data.rds")
```

MRT station names are extracted using a html function as below.

```{r}
extract_station_name <- function(html_text) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  tryCatch({
    page <- read_html(html_text)
    # Find the row that contains STATION_NA
    rows <- html_elements(page, "tr")
    
    for (row in rows) {
      th_text <- html_text(html_element(row, "th"))
      if (!is.na(th_text) && th_text == "STATION_NA") {
        station_name <- html_text(html_element(row, "td"))
        # Remove " MRT STATION" and " STATION" from the name
        station_name <- gsub(" MRT STATION$", "", station_name)
        station_name <- gsub(" LRT STATION$", "", station_name)
        return(station_name)
      }
    }
    return(NA_character_)
  }, error = function(e) {
    return(NA_character_)
  })
}
```

Station names are extracted and duplicates of station names are removed.

```{r}
# Apply to extract station names
mrt_sf$station_name <- sapply(mrt_sf$Description, extract_station_name)

mrt_sf$station_name <- sapply(mrt_sf$Description, extract_station_name) %>%
  toupper() %>%                    # Convert to uppercase
  trimws() %>%                     # Remove leading/trailing spaces
  gsub("[^A-Z0-9 ]", "", .)       # Remove special characters

# Remove duplicate stations
mrt_clean <- mrt_sf %>%
  distinct(station_name, .keep_all = TRUE)
```

```{r}
plot(st_geometry(mpsz), col = NA, border = "gray", main = "MRT Locations")
plot(st_geometry(mrt_clean), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

### 3.5 Importing Transport data

```{r}
#| echo: false
busstop <- st_read("data/geospatial/LTABusStop.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false
busstop_clean <- st_join(
  busstop, mpsz,
  join = st_within,
  left = FALSE)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Bus Stop Locations")
plot(st_geometry(busstop_clean), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

### 3.6 Importing Education data

```{r}
#| echo: false

preschool <- st_read("data/geospatial/PreSchoolsLocation.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Preschool Locations")
plot(st_geometry(preschool), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

```{r}
#| echo: false

childcare <- st_read("data/geospatial/ChildCareServices.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Childcare Locations")
plot(st_geometry(childcare), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

### 3.7 Importing Elderly Care data

```{r}
#| echo: false

eldercare <- st_read("data/geospatial/EldercareServices.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Eldercare Locations")
plot(st_geometry(eldercare), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

### 3.8 Importing Recreational data

```{r}
#| echo: false

parks <- st_read("data/geospatial/Parks.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Park Locations")
plot(st_geometry(parks), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

```{r}
#| echo: false

clubs <- st_read("data/geospatial/CommunityClubs.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Community Club Locations")
plot(st_geometry(clubs), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

```{r}
#| echo: false

sports <- st_read("data/geospatial/SportSGSportFacilitiesGEOJSON.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Sports Facilities Locations")
plot(st_geometry(sports), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

```{r}
#| echo: false

gyms <- st_read("data/geospatial/GymsSGGEOJSON.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Gym Locations")
plot(st_geometry(gyms), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

### 3.9 Importing Food data

```{r}
#| echo: false

hawkers <- st_read("data/geospatial/HawkerCentresGEOJSON.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Hawker Centre Locations")
plot(st_geometry(hawkers), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

```{r}
#| echo: false

supermarkets <- st_read("data/geospatial/SupermarketsGEOJSON.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "Supermarket Locations")
plot(st_geometry(supermarkets), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

### 3.10 Importing Healthcare data

```{r}
#| echo: false

clinics <- st_read("data/geospatial/CHASClinics.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false

plot(st_geometry(mpsz), col = NA, border = "gray", main = "CHAS Clinic Locations")
plot(st_geometry(clinics), add = TRUE, col = "black", pch = 16, cex = 0.1)
```

## 4 Data Wrangling

### 4.1 Creating Proximity and Count features

We will calculate 2 types of features for each HDB resale flat:

-   **Proximity features:** Calculate distance to nearest amenity
-   **Count features:** Count number of amenities within a range of radius

```{r}
#| eval: false

calculate_custom_spatial_features <- function(hdb_sf, 
                                             amenities_list,
                                             proximity_amenities = NULL,
                                             count_amenities = NULL,
                                             count_radii = c(350, 1000)) {
  
  # Use all amenities if none specified
  if (is.null(proximity_amenities)) {
    proximity_amenities <- names(amenities_list)
  }
  if (is.null(count_amenities)) {
    count_amenities <- names(amenities_list)
  }
  
  result <- hdb_sf
  
  # 1. CALCULATE PROXIMITY FEATURES
  cat("Calculating proximity features...\n")
  for (amenity_name in proximity_amenities) {
    tryCatch({
      amenity_sf <- amenities_list[[amenity_name]]
      prox_col <- paste0("PROX_", toupper(amenity_name))
      
      # Calculate distance to nearest feature
      nearest_indices <- st_nearest_feature(result, amenity_sf)
      distances <- st_distance(result, amenity_sf[nearest_indices, ], by_element = TRUE)
      
      result[[prox_col]] <- as.numeric(distances)
      cat("  ✓", prox_col, "\n")
      
    }, error = function(e) {
      cat("  ✗ Failed:", amenity_name, "-", e$message, "\n")
      result[[paste0("PROX_", toupper(amenity_name))]] <- NA_real_
    })
  }
  
  # 2. CALCULATE COUNT FEATURES
  cat("Calculating count features...\n")
  for (amenity_name in count_amenities) {
    tryCatch({
      amenity_sf <- amenities_list[[amenity_name]]
      
      for (radius in count_radii) {
        count_col <- paste0("COUNT_", radius, "M_", toupper(amenity_name))
        
        # Create buffers and count intersections
        buffers <- st_buffer(result, radius)
        counts <- lengths(st_intersects(buffers, amenity_sf))
        
        result[[count_col]] <- counts
        cat("  ✓", count_col, "\n")
      }
      
    }, error = function(e) {
      cat("  ✗ Failed counts for:", amenity_name, "-", e$message, "\n")
      for (radius in count_radii) {
        result[[paste0("COUNT_", radius, "M_", toupper(amenity_name))]] <- NA_real_
      }
    })
  }
  
  return(result)
}
```

```{r}
#| eval: false

# Define your amenities 
amenities <- list(
  busstop_2 = busstop_clean,
  childcare = childcare,
  clinics = clinics,
  clubs = clubs,
  eldercare = eldercare,
  gyms = gyms,
  hawkers = hawkers,
  mrt_2 = mrt_clean,
  parks = parks,
  preschool = preschool,
  schools_2 = schools_sf,
  sports = sports,
  supermarkets = supermarkets
)

# Calculate ALL proximity and count features
hdb_all_features <- calculate_custom_spatial_features(
  hdb_sf = HDBresale_sf,
  amenities_list = amenities,
  proximity_amenities = names(amenities),  # All for proximity
  count_amenities = names(amenities),      # All for counts
  count_radii = c(350, 1000)               # Use 2 radii
)
```

```{r}
#| eval: false

saveRDS(hdb_all_features, file = "data/rds/hdb_all_features.rds")
```

```{r}
hdb_all_features <- readRDS("data/rds/hdb_all_features.rds")
```

```{r}
glimpse(hdb_all_features)
```

### 4.2 Calculating Spatial features

We convert spatial features *storey_range* and *remaining_lease* into numerical columns so that it can be used for predictive modelling.

We then remove features that are non-numeric.

```{r}

# Calculate ordered column for storey_range

hdb_all_features <- hdb_all_features %>%
  mutate(
    storey_order = case_when(
      storey_range == "01 TO 03" ~ 1,
      storey_range == "04 TO 06" ~ 2,
      storey_range == "07 TO 09" ~ 3,
      storey_range == "10 TO 12" ~ 4,
      storey_range == "13 TO 15" ~ 5,
      storey_range == "16 TO 18" ~ 6,
      storey_range == "19 TO 21" ~ 7,
      storey_range == "22 TO 24" ~ 8,
      storey_range == "25 TO 27" ~ 9,
      storey_range == "28 TO 30" ~ 10,
      storey_range == "31 TO 33" ~ 11,
      storey_range == "34 TO 36" ~ 12,
      storey_range == "37 TO 39" ~ 13,
      storey_range == "40 TO 42" ~ 14,
      storey_range == "43 TO 45" ~ 15,
      storey_range == "49 TO 51" ~ 16, 
      TRUE ~ NA_real_
    )
  )

# Convert lease_remaining to months

convert_lease_to_months <- function(lease_string) {
  years <- as.numeric(str_extract(lease_string, "\\d+(?= years?)"))
  months <- as.numeric(str_extract(lease_string, "\\d+(?= months?)"))
  
  # Handle NAs
  years <- ifelse(is.na(years), 0, years)
  months <- ifelse(is.na(months), 0, months)
  
  return((years * 12) + months)
}

hdb_all_features <- hdb_all_features %>%
  mutate(remaining_lease_months = convert_lease_to_months(remaining_lease))

hdb_numeric <- hdb_all_features %>%
  select(-month, -town, -flat_type, -block, -street_name, -storey_range, -flat_model, -lease_commence_date, -remaining_lease)
```

## 5 Exploratory Data Analysis (EDA)

We plot the features and dependent variable, *Resale_Price*, using histograms. This is to understand the underlying distribution of each features.

```{r}
# Create histograms
hdb_eda <- hdb_numeric 

# Set up plotting parameters
par(mfrow = c(4, 4))  # Adjust grid size based on number of features
par(mar = c(3, 3, 2, 1))  # Adjust margins

# Create histograms for all numeric columns
for(col in names(hdb_eda)) {
  if(is.numeric(hdb_eda[[col]])) {
    hist(hdb_numeric[[col]], 
         main = col,
         xlab = "",
         ylab = "",
         col = "lightblue",
         border = "black",
         breaks = 30)
  }
}

# Reset plotting parameters
par(mfrow = c(1, 1))
par(mar = c(5, 4, 4, 2) + 0.1)  # Reset to default margins
```

From the plots we can observe that:

-   Most COUNT variables have right skew, hence log transformation is required.

-   Some COUNT variables like `COUNT_350M_SPORTS`, `COUNT_350M_GYMS`, `COUNT_350M_CLUBS` have most values at 0 and 1. We can convert to binary attributes instead.

-   Most PROX variables can be inverse transformed using *1 / (1 + df\[\[feature\]\])*, which will place more weight on closer amenities than amenities that are located further away, with a decay effect.

-   `remaining_lease_months` has a left skew, as most HDB flats in Singapore have substantial lease remaining, with fewer approaching the end of their 99-year lease. We can reflect the data, and add log transformation.

-   The dependent variable, `resale_price` has a right skew, hence log transformation is required.

```{r}
colnames(hdb_numeric)
```

We then define a feature engineering function to apply the right transformations and standardisations to each variable.

```{r}
#| eval: false

feature_engineering <- function(data) {
  df_engineered <- data
  
  # Define variable groups - FIXED: removed duplicates and inconsistencies
  count_vars_right_skew <- c(
    "COUNT_1000M_BUSSTOP_2", "COUNT_1000M_CHILDCARE", "COUNT_350M_CHILDCARE",
    "COUNT_1000M_CLINICS", "COUNT_1000M_CLUBS", "COUNT_1000M_ELDERCARE",
    "COUNT_1000M_GYMS", "COUNT_1000M_HAWKERS", "COUNT_1000M_MRT_2", 
    "COUNT_1000M_PARKS", "COUNT_1000M_PRESCHOOL", "COUNT_1000M_SCHOOLS_2",
    "COUNT_1000M_SPORTS", "COUNT_1000M_SUPERMARKETS"
  )
  
  count_vars_binary <- c(
    "COUNT_350M_BUSSTOP_2", "COUNT_350M_CLINICS", "COUNT_350M_CLUBS", 
    "COUNT_350M_ELDERCARE", "COUNT_350M_GYMS", "COUNT_350M_HAWKERS",
    "COUNT_350M_MRT_2", "COUNT_350M_PARKS", "COUNT_350M_PRESCHOOL",
    "COUNT_350M_SCHOOLS_2", "COUNT_350M_SPORTS", "COUNT_350M_SUPERMARKETS"
  )
  
  prox_vars <- c(
    "PROX_BUSSTOP_2", "PROX_CHILDCARE", "PROX_CLINICS", "PROX_CLUBS", 
    "PROX_ELDERCARE", "PROX_GYMS", "PROX_HAWKERS", "PROX_MRT_2", 
    "PROX_PARKS", "PROX_PRESCHOOL", "PROX_SCHOOLS_2", "PROX_SPORTS", 
    "PROX_SUPERMARKETS"
  )
  
  continuous_vars <- c("floor_area_sqm", "storey_order")
  
  # 1. COUNT: Right-skewed → Log transform → Standardise
  for(feature in count_vars_right_skew) {
    if(feature %in% names(data)) {
      log_transformed <- log1p(data[[feature]])
      df_engineered[[paste0("STD_LOG_", feature)]] <- as.numeric(scale(log_transformed))
      df_engineered[[feature]] <- NULL
    }
  }
  
  # 2. COUNT: Binary distribution → Keep as 0/1 (NO standardisation)
  for(feature in count_vars_binary) {
    if(feature %in% names(data)) {
      df_engineered[[paste0("BIN_", feature)]] <- as.numeric(data[[feature]] > 0)
      df_engineered[[feature]] <- NULL
    }
  }
  
  # 3. PROX: Inverse transform → Standardise
  for(feature in prox_vars) {
    if(feature %in% names(data)) {
      inv_transformed <- 1 / (1 + data[[feature]])
      df_engineered[[paste0("STD_INV_", feature)]] <- as.numeric(scale(inv_transformed))
      df_engineered[[feature]] <- NULL
    }
  }
  
  # 4. Continuous variables → Standardise
  for(feature in continuous_vars) {
    if(feature %in% names(data)) {
      df_engineered[[paste0("STD_", feature)]] <- as.numeric(scale(df_engineered[[feature]]))
      df_engineered[[feature]] <- NULL
    }
  }
  
  # 5. Remaining lease: Special handling
  if("remaining_lease_months" %in% names(data)) {
    max_lease <- max(data$remaining_lease_months, na.rm = TRUE)
    reflected_log <- log1p(max_lease - data$remaining_lease_months)
    df_engineered$STD_LOG_REFLECT_LEASE <- as.numeric(scale(reflected_log))
    df_engineered$resale_price <- NULL
  }
  
  # 6. Log transform the target variable AND keep original
  if("resale_price" %in% names(data)) {
    df_engineered$log_resale_price <- log(data$resale_price)
    # Drop the original resale_price
    df_engineered$remaining_lease_months <- NULL
  }
  
  return(df_engineered) 
}  

# Apply feature engineering
hdb_engineered <- feature_engineering(hdb_numeric)

# Check the results
glimpse(hdb_engineered)
cat("Number of features after engineering:", ncol(hdb_engineered), "\n")
cat("Target variables:", names(hdb_engineered)[grepl("resale_price", names(hdb_engineered))], "\n")
```

```{r}
#| eval: false

saveRDS(hdb_engineered, file = "data/rds/HDBresale_eng.rds")
```

```{r}
hdb_engineered <- read_rds("data/rds/HDBresale_eng.rds")
```

## 6 Computing Correlation Matrix

We examine if multicollinearity is present in the features.

```{r}
hdb_engineered_nogeo <- hdb_engineered %>%
  st_drop_geometry() %>%
  select(-log_resale_price)

corrplot::corrplot(cor(hdb_engineered_nogeo, use = "complete.obs"), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper",
                   title = "Feature Correlations (Without Price)")
```

```{r}
get_high_correlations <- function(cor_matrix, threshold = 0.8) {
  cor_matrix[lower.tri(cor_matrix, diag = TRUE)] <- NA
  cor_df <- as.data.frame(as.table(cor_matrix))
  colnames(cor_df) <- c("Feature1", "Feature2", "Correlation")
  high_corr <- cor_df %>%
    filter(abs(Correlation) > threshold & !is.na(Correlation)) %>%
    arrange(desc(abs(Correlation)))
  return(high_corr)
}

cor_matrix <- cor(hdb_engineered_nogeo, use = "complete.obs")
high_corr_pairs <- get_high_correlations(cor_matrix, threshold = 0.8)

print("Highly correlated feature pairs (r > 0.8):")

high_corr_pairs %>% 
  as_tibble() %>% 
  print(n = Inf)
```

Childcare and Preschool columns are extremely correlated with each other, hence Childcare columns will be dropped.

```{r}
hdb_final <- hdb_engineered %>%
  select(-STD_INV_PROX_CHILDCARE, -STD_LOG_COUNT_1000M_CHILDCARE)
```

```{r}
write_rds(hdb_final, "data/model/hdb_final.rds")
```

```{r}
hdb_final <- read_rds("data/model/hdb_final.rds")
```

## 7 Building a Non-spatial Multiple Linear Regression (MLR)

This will be our base model to identify important features, and to check residuals for spatial autocorrelation.

### 7.1 Data Train-test Split

The data is split into training and test data sets, 70% and 30% respectively, by using `initial_split()` of **rsample** package.

```{r}
set.seed(1234)
hdb_sampled <- hdb_final %>%
  sample_n(5000)

resale_split <- initial_split(hdb_sampled, prop = 0.7)  # 70% train, 30% test

train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
write_rds(train_data, "data/model/train_data.rds")
write_rds(test_data, "data/model/test_data.rds")
```

```{r}
train_data <- read_rds("data/model/train_data.rds")
test_data <- read_rds("data/model/test_data.rds")
```

### 7.2 Create Multiple Linear Regression model

```{r}
price_mlr <- lm(log_resale_price ~ ., 
                data = train_data %>% st_drop_geometry())
summary(price_mlr)
```

The adjusted R_squared = 0.661, which means that the features explain 66.1% of variability in HDB resale price. The residual SE = 0.123 which means approximately 68% of predictions will be within 88.4% to 113.1% of actual price.

The p-value is \< 0.05, indicating that the model is statistically significant, meaning the predictors collectively contribute significantly to predicting the resale price.

### 7.3 Retain Important Features

First, we identify the features that are statistically significant (i.e. p \< 0.05).

```{r}
important_vars <- names(which(summary(price_mlr)$coefficients[-1, 4] < 0.05))
cat("Statistically significant predictors:", length(important_vars), "\n")
```

```{r}
coef_table <- as.data.frame(summary(price_mlr)$coefficients) %>%
  rownames_to_column(var = "Predictor")

labeled_features <- coef_table %>%
  rename(P_Value = `Pr(>|t|)`) %>%
  filter(Predictor != "(Intercept)") %>%
  mutate(
    Significance = ifelse(P_Value < 0.05, "Significant", "Non-Significant")
  ) %>%
  select(Predictor, Estimate, P_Value, Significance)
print(labeled_features)
```

Next, we filter the train and test data to remove columns that are insignificant features.

```{r}
# Extract significant predictors (p < 0.05)
significant_features <- labeled_features %>%
  filter(Significance == "Significant") %>%
  pull(Predictor)

print("Significant features to keep:")
print(significant_features)
```

```{r}
# Keep only significant features + the target variable + geometry
train_data_filtered <- train_data %>%
  select(log_resale_price, all_of(significant_features), geometry)

cat("Training data dimensions after filtering:", dim(train_data_filtered), "\n")

# Keep same features in test data
test_data_filtered <- test_data %>%
  select(log_resale_price, all_of(significant_features), geometry)

cat("Test data dimensions after filtering:", dim(test_data_filtered), "\n")
```

```{r}
# Check if both datasets have same columns
cat("Training columns:", names(train_data_filtered), "\n")
cat("Test columns:", names(test_data_filtered), "\n")
cat("Columns match:", identical(names(train_data_filtered), names(test_data_filtered)), "\n")
```

```{r}
price_mlr_filtered <- lm(log_resale_price ~ ., 
                data = train_data_filtered %>% st_drop_geometry())
summary(price_mlr_filtered)
```

After removing 11 features:

The adjusted R_squared = 0.661, which is at a similar level as the previous model. The residual SE = 0.123 is similar level as the previous model.

The p-value is \< 0.05, indicating that the model is statistically significant, meaning the predictors collectively contribute significantly to predicting the resale price.

Residual SE and adjusted R_squared remains the same and the model will experience less overfitting with fewer features.

### 7.4 Check MLR Residuals for Spatial Autocorrelation

```{r}
#| eval: false

# Get the exact same data that was used in the model
model_indices <- as.numeric(rownames(model.frame(price_mlr_filtered)))
train_data_used_in_model <- train_data_filtered[model_indices, ]

# Add residuals
train_data_used_in_model$mlr_residuals <- residuals(price_mlr_filtered)

# Perform Moran's I on the same data the model used
moran_test <- moran.mc(train_data_used_in_model$mlr_residuals, 
                       listw = nb2listw(knn2nb(knearneigh(st_centroid(train_data_used_in_model)))), 
                       nsim = 999)

cat("Moran's I p-value:", moran_test$p.value, "\n")
cat("Number of observations used:", nrow(train_data_used_in_model), "\n")

# Save results
saveRDS(moran_test, "data/model/moran_test_filtered.rds")
```

```{r}
#| echo: false

moran_test_filtered <- readRDS("data/model/moran_test_filtered.rds")
print(moran_test_filtered)
```

Moran's I test on MLR residuals revealed significant spatial autocorrelation (p \< 0.05), indicating that housing prices vary geographically across Singapore. This strongly justifies the use of spatial modeling approaches like GWR and GWRF to capture these location-specific relationship which we will explore in the following sections.

## 8 Building a Geographically Weighted Regression Model (GWR)

From Moran's I test, we ascertain that MLR residuals show spatial patterns. Hence, we build a GWR model to capture spatial heterogeneity as feature relationships with price may vary by location.

### 8.1 Computing Adaptive Bandwidth

Firstly, `bw.gwr()` of **GWmodel** package will be used to determine the optimal bandwidth to be used.

```{r}
predictors <- names(train_data_filtered)
predictors <- predictors[!predictors %in% c("log_resale_price", "geometry")]

gwr_formula <- as.formula(paste("log_resale_price ~", paste(predictors, collapse = " + ")))
```

```{r}
#| eval: false

bw_adaptive <- bw.gwr(
  formula = gwr_formula,  
  data = train_data_filtered,  
  approach = "AIC",
  kernel = "gaussian",
  adaptive = TRUE,
  longlat = FALSE
)

write_rds(bw_adaptive, "data/model/bw_adaptive.rds")
```

```{r}
bw_adaptive <- read_rds("data/model/bw_adaptive.rds")

bw_adaptive
```

The optimal adaptive bandwidth of 74 neighbors was determined using AIC criterion, indicating localised spatial effects in Singapore's housing market where each regression uses the 74 closest transactions.

### 8.2 Constructing the Adaptive Bandwidth GWR Model

The GWR model was calibrated with the optimal bandwidth, generating location-specific coefficients that capture how amenity effects vary geographically across Singapore.

```{r}
#| eval: false

gwr_adaptive <- gwr.basic(
  formula = gwr_formula,  
  data = train_data_filtered,
  bw = bw_adaptive, 
  kernel = 'gaussian', 
  adaptive = TRUE,
  longlat = FALSE
)

write_rds(gwr_adaptive, "data/model/gwr_adaptive.rds")
```

```{r}
gwr_adaptive <- read_rds("data/model/gwr_adaptive.rds")

gwr_adaptive
```

### 8.3 Computing Adaptive Bandwidth for the Test Data

The bandwidth calculation is repeated, specifically using the coordinates of the test data locations, to compare against train data bandwidth calculations.

```{r}
#| eval: false

predictors <- names(test_data)
predictors <- predictors[!predictors %in% c("log_resale_price", "geometry")]

gwr_formula_test <- as.formula(paste("log_resale_price ~", paste(predictors, collapse = " + ")))

gwr_bw_test_adaptive <- bw.gwr(
  formula = gwr_formula,  
  data = test_data_filtered,
  approach = "AIC",
  kernel = "gaussian",
  adaptive = TRUE,
  longlat = FALSE
)

write_rds(gwr_bw_test_adaptive, "data/model/gwr_bw_test_adaptive.rds")
```

```{r}
gwr_bw_test_adaptive <- read_rds("data/model/gwr_bw_test_adaptive.rds")

gwr_bw_test_adaptive
```

For test data sample, the optimal adaptive bandwidth of 32 neighbors was determined using AIC criterion. However since this differs greatly from the training adaptive bandwidth (74 neighbours), we will use the training adaptive bandwidth.

### 8.4 Computing Predicted Values of the Test Data

We then generate the predictions from test data.

```{r}
#| eval: FALSE

train_sp_filtered <- as_Spatial(train_data_filtered)
test_sp_filtered  <- as_Spatial(test_data_filtered)

gwr_pred <- gwr.predict(
  formula = gwr_formula, 
  data = train_sp_filtered,          
  predictdata = test_sp_filtered,    
  bw = bw_adaptive, 
  kernel = 'gaussian', 
  adaptive = TRUE, 
  longlat = FALSE)

write_rds(gwr_pred, 
          "data/model/gwr_pred.rds")
```

```{r}
gwr_pred <- read_rds("data/model/gwr_pred.rds")

gwr_pred
```

## 9 Building a Random Forest Model

Random Forest is a non-parametric ensemble method that is robust to multicollinearity and automatically handles complex, non-linear relationships. We use it here as a high-performance predictive benchmark against the GWR model.

We will calibrate a Random Forest model to predict HDB resale price by using `random forest` function of **ranger** package.

### 9.1 Preparing Data for Non-Spatial RF

Since Random Forest does not use geographic coordinates for its core calculations, we strip the *geometry* column from the filtered training and test datasets using `st_drop_geometry()` of **sf** package.

```{r}
train_rf <- train_data_filtered %>% 
  st_drop_geometry()

test_rf <- test_data_filtered %>% 
  st_drop_geometry()
```

### 9.2 Constructing the Random Forest Model

We use the **ranger** package to fit the RF model.

```{r}
#| eval: false

predictors <- names(train_data_filtered)
predictors <- predictors[!predictors %in% c("log_resale_price", "geometry")]

gwr_formula <- as.formula(paste("log_resale_price ~", paste(predictors, collapse = " + ")))

# The RF model will be trained on the log-transformed price
set.seed(42) 
rf_model <- ranger::ranger(
  formula = gwr_formula,              
  data = train_rf,                    
  num.trees = 500,                    
  importance = "permutation",         
  seed = 42                           
)

write_rds(rf_model, "data/model/rf_model.rds")
```

```{r}
rf_model <- read_rds("data/model/rf_model.rds")

rf_model
```

### 9.3 Computing Predicted Values of the Test Data

We then generate the predictions from test data.

```{r}
#| eval: FALSE

rf_pred <- predict(rf_model, data = test_rf)

# The prediction object contains the predicted values in the 'predictions' slot
rf_pred_values <- rf_pred$predictions

write_rds(rf_pred_values, "data/model/rf_pred.rds")
```

```{r}
rf_pred <- read_rds("data/model/rf_pred.rds")

rf_pred
```

## 10 Building a Geographically Weighted Random Forest Model (GWRF)

GWRF, using the `grf()` function from the SpatialML package, combines the predictive power of Random Forest with local spatial modeling. It is expected to improve performance by accounting for geographical variations in feature relationships without relying on parametric assumptions.

We will calibrate a model to predict HDB resale price by using `grf()` of **SpatialML** package.

### 10.1 Preparing Spatial Data for GWRF

The `grf()` function does not accept sf or sp objects directly; it requires a standard data frame (dframe) and separate coordinate columns (coords). We must extract the coordinates from our filtered spatial data.

```{r}
# 1. Extract Coordinates from filtered data
coords_train <- st_coordinates(train_data_filtered)
coords_test  <- st_coordinates(test_data_filtered)

# 2. Create non-spatial data frames for the 'dframe' argument,
train_gwrf <- train_data_filtered %>%
  st_drop_geometry()

test_gwrf <- test_data_filtered %>%
  st_drop_geometry()

# Check dimensions
cat("GWRF Training data dimensions:", dim(train_gwrf), "\n")
cat("GWRF Test data dimensions:", dim(test_gwrf), "\n")
```

