---
title: "3: Modelling HDB Resale Prices with Geographically Weighted Methods"
author: "Emma Chew En Chin"
date: "Oct 26, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

In Singapore, understanding the factors that influence the potential prices of HDB resale flats is crucial. For prospective homeowners, this knowledge is vital as it directly impacts the long-term financial sustainability of the purchase, including the amount of housing equity they can secure and how well the flat retains its value over time. This understanding allows buyers to make informed and prudent housing decisions regarding the flat size and location that best meet their family's needs and financial standing.

This study utilises **Geographically Weighted Models (GWMs)** to analyse the key factors that influence HDB resale prices, specifically for **4-room flats** due to its popularity with young families. From this, we can produce local coefficients for each data point (or small region), reflecting the spatial heterogeneity in the data. The data period for analysis will be 1 January 2025 to 30 September 2025.

## 2 Installing and Loading R packages

**Data Acquisition & Handling** - `httr`: Maakes HTTP requests to APIs; used for url call from OneMap. - `jsonlite`: Parses JSON data returned from web APIs into usable R data frames. - `tidyverse`: A meta-package for data manipulation (dplyr), visualization (ggplot2), and more. The core toolkit for data wrangling.

**Geospatial Core** - `sf`: Handles spatial data (points, polygons). Used for spatial joins, distance calculations, and mapping. - `tmap`: Creates thematic maps (chloropleth, point maps) to visualise spatial patterns in prices and model results.

**Geospatial Modelling & Analysis** - `spdep`: Provides spatial dependence tests (e.g., Moran's I for spatial autocorrelation) and classic spatial econometric models. - `GWmodel`: The primary package for GWR and multiscale GWR. Fits models where coefficients vary geographically. - `SpatialML`: Provides geographically weighted random forests (gwrf()), blending random forests with local modeling.

**Model Pipeline and Evaluation** - `rsample`: Creates training/test splits for model validation using techniques like cross-validation. - `Metrics`: Calculates model performance metrics (RMSE, MAE, R²) to evaluate prediction accuracy.

**Utility** - `progress`: Adds progress bars to loops and long-running operations, improving user experience during spatial computations.

```{r}
pacman::p_load(httr, sf, tmap, jsonlite, progress,
               spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse)
```

## 3 Data Preparation

### 3.1 The Datasets

**Aspatial dataset:**

-   HDB Resale data: a list of HDB resale transacted prices in Singapore from Jan 2017 onwards.

**Geospatial dataset:**

-   MP14_SUBZONE_WEB_PL: a polygon feature data providing information of URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format.

-   Transport Factors:

    -   *MRT* data is a list of MRT/LRT stations in Singapore with the station names and codes. It is in geojson format.
    -   *Bus stops* data is a list of bus stops in Singapore. It is in geojson format.

-   Education Factors:

    -   *Preschool* data is a list of preschools in Singapore. It is in geojson format.
    -   *Childcare service* data is a list of childcare services in Singapore. It is in geojson format.
    -   *School* data is a list of primary, secondary, junior colleges schools in Singapore. It is in csv format.

-   Elderly-care Factors:

    -   *Eldercare* data is a list of eldercare in Singapore. It is in geojson format.

-   Recreational Factors

    -   *Parks* data is a list of parks in Singapore. It is in geojson format.
    -   *Community clubs* data is a list of community clubs in Singapore. It is in geojson format.
    -   *Community clubs* data is a list of community clubs in Singapore. It is in geojson format.
    -   *Sports Facilities* data is a list of public sports facilities in Singapore. It is in geojson format.
    -   *GymsSG* data is a list of public gym facilities in Singapore. It is in geojson format.

-   Food Factors

    -   *Hawker Centre* data is a list of hawker centres in Singapore. It is in geojson format.
    -   *Supermarket* data is a list of supermarkets in Singapore. It is in geojson format.

-   Healthcare Factors

    -   *CHAS clinics* data is a list of CHAS clinics in Singapore. It is in geojson format.

### 3.2 Cleaning HDB Resale Data

We can reverse geocode the HDB postal code, as there is only blk number and street name in the Resale Flats dataset.

This is the documentation: [OneMap Reverse Geocoding](https://www.onemap.gov.sg/apidocs/reverseGeocode)

```{r}
HDBresale <-read_csv("data/rawdata/HDBresale.csv") %>%
  filter(flat_type == "4 ROOM") %>%
  filter(month >= "2025-01" & month < "2025-09")

glimpse(HDBresale)
```

Changing ST to SAINT so that it does not look like 'Street'.

```{r}
HDBresale$street_name <- gsub(
  "ST\\.", "SAINT",
  HDBresale$street_name)
```

```{r}
geocode <- function(block, streetname) {
  base_url <- "https://onemap.gov.sg/api/common/elastic/search"
  address <- paste(block, streetname, sep = " ")
  query <- list("searchVal" = address,
                "returnGeom" = "Y",
                "getAddrDetails" = "N",
                "pageNum" = "1")
  
  res <-GET(base_url, query = query)
  restext <- content(res, as="text")
  
  output <- fromJSON(restext) %>%
    as.data.frame %>%
    select(results.LATITUDE, results.LONGITUDE)
  
  return(output)
}
```

```{r}
#| eval: false
#| warning: false
#| results: 'hide'
#| message: false

HDBresale$LATITUDE <- 0
HDBresale$LONGITUDE <- 0

for (i in 1:nrow(HDBresale)){
  temp_output <- geocode(HDBresale[i, 4], HDBresale[i, 5])
  
  HDBresale$LATITUDE[i] <- temp_output$results.LATITUDE
  HDBresale$LONGITUDE[i] <- temp_output$results.LONGITUDE
}
```

```{r}
glimpse(HDBresale)
```

Since HDBresale contains LATITUDE and LONGITUDE columns (which are in the WGS 84 geographic coordinate system, or EPSG:4326), we need to convert it into an sf (Simple Features) object and then transform it to the SVY21 (Singapore Projected Coordinate System, EPSG:3414).

```{r}
HDBresale_sf <- HDBresale %>%
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs=4326) %>%
  st_transform(crs = 3414)

saveRDS(HDBresale_sf, file = "data/rds/HDBresale_sf.rds")
```

### 3.3 Cleaning Schools Data

```{r}
folder_path <- "data/aspatial/acra"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$", 
                        full.names = TRUE)
acra_data <- file_list %>%
    map_dfr(read_csv, show_col_types = FALSE)
```
