---
title: "5: Global and Local Measures of Spatial Autocorrelation: sfdep methods"
author: "Emma Chew En Chin"
date: "Sep 27, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Getting Started

Loading all the R packages

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, sfdep)
```

## 2 Importing the Data

### 2.1 Importing the Shapefile

Hunan geospatial data is projected from from WGS 84 to UTM zone 50N (i.e. EPSG: 32650).

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan") %>% 
  st_transform(crs = 32650)
```

### 2.2 Importing the csv file

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### 2.3 Performing a Relational Join

The attribute table of *hunan*'s SpatialPolygonsDataFrame will be updated with the attribute fields of *hunan2012* dataframe.

```{r}
hunan_GDPPC <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```

### 2.4 Plotting a Choropleth Map

```{r}
#| fig-width: 9
#| fig-height: 7
#| fig-align: left
#| echo: false
tmap_mode("plot")
tm_shape(hunan_GDPPC) +
  tm_fill("GDPPC", 
          style = "quantile", 
          palette = "Blues",
          title = "GDPPC") +
  tm_layout(main.title = "Distribution of GDP per capita by county, Hunan Province",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2)
```

## 3 Global Measures of Spatial Association

Step 1: Deriving Queen's contiguity weights: sfdep methods

```{r}
wm_q <- hunan_GDPPC %>%
    mutate(nb = st_contiguity(geometry),
           wt = st_weights(nb, 
                          style = "W"),
           .before = 1)
```


### 3.1 Computing Global Moran' I

```{r}
moranI <- global_moran(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)
glimpse(moranI)
```

### 3.2 Performing Global Moran's I test

```{r}
global_moran_test(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)

```

### 3.3 Performing Global Moran'I permutation test

Step 1: use `set.seed()` before performing simulation to ensure code replication

```{r}
set.seed(1234)
```

Step 2: use `global_moran_perm()` to perform Monte Carlo simulation.

```{r}
global_moran_perm(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt,
                  nsim = 99)
```

The p-value < 0.01, hence we can reject the null hypothesis that the points are arranged randomly at a 99% confidence interval. Additionally the Moran's I value is 0.301, which indicates a moderate level of clustering present.

## 4 Local Measures of Spatial Association

### 4.1 Visualise Local Moran's I

```{r}
lisa <- wm_q %>% 
  mutate(local_moran = local_moran(
    GDPPC, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)

#| fig-width: 8
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(
    main.title = "local Moran's I of GDPPC",
    main.title.size = 2)
```

### 4.2 Visualising p-value of local Moran's I

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("p_ii_sim") + 
  tm_borders(alpha = 0.5) +
   tm_layout(main.title = "p-value of local Moran's I",
            main.title.size = 2)
```
### 4.3 Visualising Local Moran's I and p-values

```{r}
#| fig-height: 7
tmap_mode("plot")
map1 <- tm_shape(lisa) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "local Moran's I of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(lisa) +
  tm_fill("p_ii",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of local Moran's I",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

### 4.4 Plot LISA map (quadrants)

```{r}
lisa_sig <- lisa  %>%
  filter(p_ii_sim < 0.05)
tmap_mode("plot")
tm_shape(lisa) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(lisa_sig) +
  tm_fill("mean") + 
  tm_borders(alpha = 0.4)
```

## 5 Hot Spot and Cold Spot Area Analysis (HCSA)

Deriving special weight matrix based on distances.

```{r}
wm_idw <- hunan_GDPPC %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wts = st_inverse_distance(nb, 
                              geometry, 
                              scale = 1,
                              alpha = 1),
         .before = 1)
```

Compute local Gi* statistics.

```{r}
HCSA <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    GDPPC, nb, wts, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
HCSA
```

### 5.1 Visualising Gi*

```{r}
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8))
```

### 5.2 Visualising p-value of HCSA

```{r}
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("p_sim") + 
  tm_borders(alpha = 0.5)
```

### 5.3 Visualising local HCSA

```{r}
tmap_mode("plot")
map1 <- tm_shape(HCSA) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "Gi* of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(HCSA) +
  tm_fill("p_value",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of Gi*",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

### 5.4 Visualising hot spot and cold spot areas

```{r}
HCSA_sig <- HCSA  %>%
  filter(p_sim < 0.05)
tmap_mode("plot")
tm_shape(HCSA) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(HCSA_sig) +
  tm_fill("cluster") + 
  tm_borders(alpha = 0.4)
```

Figure above reveals that there is one hot spot area and two cold spot areas. Interestingly, the hot spot areas coincide with the High-high cluster identifies by using local Moran's I method in the earlier sub-section.