---
title: "5: Global and Local Measures of Spatial Autocorrelation: sfdep methods"
author: "Emma Chew En Chin"
date: "Sep 27, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Getting Started

Loading all the R packages

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, sfdep)
```

## 2 Importing the Data

### 2.1 Importing the Shapefile

Hunan geospatial data is projected from from WGS 84 to UTM zone 50N (i.e. EPSG: 32650).

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan") %>% 
  st_transform(crs = 32650)
```

### 2.2 Importing the csv file


```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### 2.3 Performing a Relational Join

The attribute table of *hunan*'s SpatialPolygonsDataFrame will be updated with the attribute fields of *hunan2012* dataframe.

```{r}
hunan_GDPPC <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```
## 3 Global Measures of Spatial Association

Step 1: Deriving Queen's contiguity weights: sfdep methods

```{r}
wm_q <- hunan_GDPPC %>%
    mutate(nb = st_contiguity(geometry),
           wt = st_weights(nb, 
                          style = "W"),
           .before = 1)
```

```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
print(wm_q)
```
