---
title: "7: Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method"
author: "Emma Chew En Chin"
date: "Oct 15, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

**Geographically weighted regression (GWR)** is a spatial statistical technique that takes non-stationary variables into consideration (e.g., climate; demographic factors; physical environment characteristics) and models the local relationships between these independent variables and an outcome of interest (also known as dependent variable).

In this exercise, a hedonic pricing model is built using GWR. The dependent variable is the resale prices of condominium in 2015. The independent variables are divided into either structural and locational.

The datasets used are the following:

-   *URA Master Plan subzone boundary* in shapefile format (i.e. MP14_SUBZONE_WEB_PL)
-   *condo_resale_2015* in csv format (i.e. condo_resale_2015.csv)

## 2 Data Preparation

### 2.1 Importing R packages

The following new packages will be used:
-   `olsrr`: For building OLS and performing diagnostics tests
-   `GWmodel`: For calibrating geographical weighted family of models
-   `corrplot`: For multivariate data visualisation and analysis

The **GWmodel** package provides a collection of localised spatial statistical methods, namely: GW summary statistics, GW principal components analysis, GW discriminant analysis and various forms of GW regression; some of which are provided in basic and robust (outlier resistant) forms.

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, 
               sf, sfdep, GWmodel, tmap, 
               tidyverse, gtsummary,
               performance, RColorBrewer)
```

### 2.2 Importing Datasets

#### 2.2.1 Importing Geospatial Data 

The MP14_SUBZONE_WEB_PL shapfile is imported and the data is in SVY21 projected coordinates systems.

```{r}
mpsz = st_read(dsn = "data/geospatial", 
               layer = "MP14_SUBZONE_WEB_PL")
```

The mpsz file is transformed with the right EPSG code (3414).

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
st_crs(mpsz_svy21)
```

The extent of mpsz_svy21 is observed using `st_bbox()` of **sf** package.

```{r}
st_bbox(mpsz_svy21)
```

#### 2.2.2 Importing Aspatial Data

```{r}
condo_resale = read_csv(
  "data/aspatial/Condo_resale_2015.csv")
```

Using glimpse, we can preview the data structure of the dataset.

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE)
```

```{r}
head(condo_resale$LATITUDE)
```

From the output above, we can see that the values of LONGITITUDE and LATITUDE fields are in decimal degree. Most probably WGS84 geographic coordinate system is used.

We can then use `summary()` to display the summary statistics of the *cond_resale* tibble file.

```{r}
summary(condo_resale)
```

As the *condo_resale* tibble data frame is aspatial, we will convert it to a **sf** object.

`st_transform()` is used to convert the coordinates from WGS84 (i.e. crs:4326) to SVY21 (i.e. crs=3414).

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

We then use `head()` to view the content of the *condo_resale* sf object.

```{r}
head(condo_resale.sf)
```

## 3 Exploratory Data Analysis (EDA)

Using **ggplot2** package, we can perform EDA.

### 3.1 EDA using statistical graphics

We can plot the distribution of *SELLING_PRICE*. 

```{r}
ggplot(data=condo_resale.sf, 
       aes(x=`SELLING_PRICE`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

The graph shows a right-skewed distribution, showing that more condominium units were transacted at lower prices.

We can normalise the skewed distribution using log transformation. 

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))
```

Plotting *LOG_SELLING_PRICE* 

```{r}
ggplot(data=condo_resale.sf, 
       aes(x=`LOG_SELLING_PRICE`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

### 3.2 Multiple Histogram Plots distribution of variables

We use `ggarrange()` of *ggpubr* package to plot multiple histograms.

```{r}
AREA_SQM <- ggplot(data=condo_resale.sf, 
                   aes(x= `AREA_SQM`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

AGE <- ggplot(data=condo_resale.sf, 
              aes(x= `AGE`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_CBD <- ggplot(data=condo_resale.sf, 
                   aes(x= `PROX_CBD`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, 
                         aes(x= `PROX_CHILDCARE`)) + 
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, 
                           aes(x= `PROX_ELDERLYCARE`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_URA_GROWTH_AREA`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, 
                             aes(x= `PROX_HAWKER_MARKET`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, 
                            aes(x= `PROX_KINDERGARTEN`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_MRT <- ggplot(data=condo_resale.sf, 
                   aes(x= `PROX_MRT`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_PARK <- ggplot(data=condo_resale.sf, 
                    aes(x= `PROX_PARK`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                           aes(x= `PROX_PRIMARY_SCH`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_TOP_PRIMARY_SCH`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, 
          PROX_ELDERLYCARE, PROX_URA_GROWTH_AREA,
          PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT,
          PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,  
          ncol = 3, nrow = 4)
```

### 3.3 Drawing Statistical Point Map

We want to reveal the geospatial distribution condominium resale prices in Singapore using the **tmap** package.

```{r}
tmap_mode("view")
```

`set_zoom_limits` argument of `tm_view()` sets the minimum and maximum zoom level to 11 and 14 respectively.

```{r}
tm_shape(mpsz_svy21)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(fill = "SELLING_PRICE",
          fill_alpha = 0.6,
          size = 0.5,
          fill.scale = tm_scale_intervals(
            style="quantile")) +
  tm_view(set_zoom_limits = c(11,14))
```

We can turn R display into `plot` mode.

```{r}
tmap_mode("plot")
```

## 4 Hedonic Pricing Modelling in R

The hedonic pricing model for condominium resale units will be built using using `lm()` of R base.

### 4.1 Simple Linear Regression Method

First, we will build a simple linear regression model by using *SELLING_PRICE* as the dependent variable and *AREA_SQM* as the independent variable.

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, 
                data = condo_resale.sf)
```

In R, the `lm()` function creates a linear model object. This object contains all the results of the regression.

To view and use these results,other functions that interact with this object is required:

-   `summary()` provides detailed coefficients, R-squared, and p-values.
-   `anova()` provides an analysis of variance table.
-   Accessor functions like `coefficients()` and `residuals()` extract specific pieces of information from the model object.

```{r}
summary(condo.slr)
```

The output report reveals that the *SELLING_PRICE* can be explained by using the formula:
      *y = -258121.1 + 14719x1*
The R-squared of 0.4518 reveals that the simple regression model is able to explain about 45% of the resale prices.

Since p-value is much smaller than 0.0001, we will reject the null hypothesis that mean is a good estimator of *SELLING_PRICE*. This will allow us to infer that simple linear regression model above is a good estimator of *SELLING_PRICE*.

The Coefficients: section of the report reveals that the p-values of both the estimates of the Intercept and ARA_SQM are smaller than 0.001. In view of this, the null hypothesis of the B0 and B1 are equal to 0 will be rejected. As a results, we will be able to infer that the B0 and B1 are good parameter estimates.

We then plot the best fit line on a scatterplot.

```{r}
ggplot(data=condo_resale.sf,  
       aes(x=`AREA_SQM`, y=`SELLING_PRICE`)) +
  geom_point() +
  geom_smooth(method = lm)
```

The plot shows a few statistical outliers with relatively high selling price.

### 4.2 Multiple Linear Regression Method

### 4.2.1 Visualising the relationships of the independent variables

```{r}
```


```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```