---
title: "8: Geographically Weighted Predictive Modelling"
author: "Emma Chew En Chin"
date: "Oct 20, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

**Geospatial predictive modeling** operates on the principle that event occurrences are not random or uniform in space. It uses geographically referenced data to describe how geospatial factors (like infrastructure, topography, or sociocultural elements) constrain and influence where events occur. The models achieve this by spatially correlating historical event locations with environmental factors that represent these constraints.

## 2 The Datasets

**Aspatial dataset:**

-   HDB Resale data: a list of HDB resale transacted prices in Singapore from Jan 2017 onwards.

**Geospatial dataset:**

-   MP14_SUBZONE_WEB_PL: a polygon feature data providing information of URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format.

-   Locational factors with geographic coordinates:

    -   *Eldercare* data is a list of eldercare in Singapore. It is in shapefile format.
    -   *Hawker Centre* data is a list of hawker centres in Singapore. It is in geojson format.
    -   *Parks* data is a list of parks in Singapore. It is in geojson format.
    -   *Supermarket* data is a list of supermarkets in Singapore. It is in geojson format.
    -   *CHAS clinics* data is a list of CHAS clinics in Singapore. It is in geojson format.
    -   *Childcare service* data is a list of childcare services in Singapore. It is in geojson format.
    -   *Kindergartens* data is a list of kindergartens in Singapore. It is in geojson format.
    -   *MRT* data is a list of MRT/LRT stations in Singapore with the station names and codes. It is in shapefile format.
    -   *Bus stops* data is a list of bus stops in Singapore. It is in shapefile format.

**Locational factors without geographic coordinates:**

-   *Primary school* data is extracted from the list on General information of schools from data.gov portal. It is in csv format.

**Retrieved/Scraped from other sources**

-   *CBD coordinates* obtained from Google.
-   *Shopping malls* data is a list of Shopping malls in Singapore obtained from Wikipedia.
-   *Good primary schools* is a list of primary schools that are ordered in ranking in terms of popularity and this can be found at Local Salary Forum.

## 3 Installing and Loading R packages

The following new packages will be used:

-   `SpatialML`:Used for spatial machine learning and predictive modeling, often providing spatial extensions to common machine learning algorithms.
-   `rsample`: Used for creating data splits and resamples (e.g., cross-validation, bootstrapping) for proper training, testing, and evaluation of predictive models.
-   `metrics`: Provides various performance metrics (like RMSE, MAE, R-squared) to evaluate and compare the accuracy of predictive and machine learning models.

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse)
```

## 4 Data Preparation

### 4.1 Reading data from rds file

```{r}
mdata <- read_rds("data/rawdata/mdata.rds")
```

### 4.2 Data Sampling

The entire data is split into training and test data sets, 65% and 35% respectively, by using `initial_split()` of **rsample** package.

**rsample** is one of the package of tidymodels.

```{r}
set.seed(1234)
resale_split <- initial_split(mdata, 
                              prop = 6.5/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
write_rds(train_data, "data/model/train_data.rds")
write_rds(test_data, "data/model/test_data.rds")
```

## 5 Computing Correlation Matrix

Before we load the predictors into a predictive model, we will use a correlation matrix to examine if there is sign of multicollinearity.

```{r}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
corrplot::corrplot(cor(mdata_nogeo[, 2:17]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper")
```

From the matrix, we can observe that all the correlation values are below 0.8, hence there is no sign of multicollinearity.

## 6 Building a non-spatial multiple linear regression

First, we retrieve the train and test datasets

```{r}
train_data <- read_rds("data/model/train_data.rds")
test_data <- read_rds("data/model/test_data.rds")
```

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
summary(price_mlr)
```

The adjusted R-squared of the model is 0.737, indicating that 73.73% of the variability in the resale price can be explained by the independent variables in the model.

The Residual Standard Error (RSE) is 61,650, which suggests that the typical or average difference between the model's predicted resale price and the actual resale price is approximately \$61,650.

The p-value is \< 0.05, indicating that the model is statistically significant, meaning the predictors collectively contribute significantly to predicting the resale price.

```{r}
write_rds(price_mlr, "data/model/price_mlr.rds" ) 
```

## 7 gwr Predictive Method

We will calibrate a model to predict HDB resale price by using geographically weighted regression method of the **GWmodel** package.

### 7.1 Computing adaptive bandwidth

Firstly, `bw.gwr()` of **GWmodel** package will be used to determine the optimal bandwidth to be used.

```{r}
#| eval: FALSE

bw_adaptive <- bw.gwr(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths +
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
    WITHIN_1KM_PRISCH,
  data=train_data,
  approach="CV",
  kernel="gaussian",
  adaptive=TRUE,
  longlat=FALSE)

write_rds(bw_adaptive, "data/model/bw_adaptive.rds")
```

The result shows that 40 neighbour points will be the optimal bandwidth to be used if adaptive bandwidth is used for this data set.

Note that this is a computational intensity and time consuming process. It is wiser to save the output as an rds file for future used without having to re-run the process again.

### 7.2 Constructing the adaptive bandwidth gwr model

```{r}
bw_adaptive <- read_rds(
  "data/model/bw_adaptive.rds")
```

```{r}
#| eval: FALSE

gwr_adaptive <- gwr.basic(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
  data=train_data,
  bw=bw_adaptive, 
  kernel = 'gaussian', 
  adaptive=TRUE,
  longlat = FALSE)

write_rds(gwr_adaptive, 
          "data/model/gwr_adaptive.rds")
```

Note that this is a computational intensity and time consuming process. It is wiser to save the output as an rds file for future used without having to re-run the process again.

### 7.3 Retrieve gwr output object

```{r}
gwr_adaptive <- read_rds("data/model/gwr_adaptive.rds")
```

```{r}
gwr_adaptive
```

### 7.4 Computing adaptive bandwidth for the test data

```{r}
#| eval: FALSE

gwr_bw_test_adaptive <- bw.gwr(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths +
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
    WITHIN_1KM_PRISCH,
  data=test_data,
  approach="CV",
  kernel="gaussian",
  adaptive=TRUE,
  longlat=FALSE)

write_rds(gwr_bw_test_adaptive, 
          "data/model/gwr_bw_test_adaptive.rds")
```

### 7.5 Computing predicted values of the test data

```{r}
#| eval: FALSE
train_data <- read_rds("data/model/train_data.rds")
test_data <- read_rds("data/model/test_data.rds")

train_sp <- as_Spatial(train_data)
test_sp  <- as_Spatial(test_data)

gwr_pred <- gwr.predict(
  formula = resale_price ~ floor_area_sqm + 
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER + 
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
    WITHIN_1KM_PRISCH, 
  data = train_sp,          
  predictdata = test_sp,    
  bw = 40, 
  kernel = 'gaussian', 
  adaptive = TRUE, 
  longlat = FALSE)

write_rds(gwr_pred, 
          "data/model/gwr_pred.rds")
```

## 8 Calibrating Random Forest Model

We will calibrate a model to predict HDB resale price by using `random forest` function of **ranger** package.

### 8.1 Preparing coordinates data

We extract the x,y coordinates of the full, training and test data sets.

```{r}
coords <- st_coordinates(mdata)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

```{r}
coords_train <- write_rds(
  coords_train, 
  "data/model/coords_train.rds" )
coords_test <- write_rds(
  coords_test, 
  "data/model/coords_test.rds" )
```

### 8.2 Dropping geometry field

First, we will drop geometry column of the sf data.frame by using `st_drop_geometry()` of **sf** package.

```{r}
train_data_nogeom <- train_data %>% 
  st_drop_geometry()

test_data_nogeom <- test_data %>% 
  st_drop_geometry()
```

### 8.3 Calibrating a non-spatial random forest model

```{r}
set.seed(1234)
rf <- ranger(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN + 
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS +  
    WITHIN_1KM_PRISCH,
  data=train_data_nogeom)
rf
```

## 9 Calibrating Geographical Weighted Random Forest Model

We will calibrate a model to predict HDB resale price by using `grf()` of **SpatialML** package.

### 9.1 Calibrating using training data

```{r}
#| eval: FALSE

set.seed(1234)
gwRF_adaptive <- grf(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
  dframe=train_data_nogeom, 
  bw=40,
  kernel="adaptive",
  coords = coords_train,
  ntree = 50,  # Reduce number of trees
  mtry = 4,     # Set mtry parameter
  importance = "impurity")

write_rds(gwRF_adaptive, "data/model/gwRF_adaptive.rds")
```

### 9.2 Predicting by using test data

#### 9.2.1 Preparing the test data

```{r}
gwRF_adaptive <- read_rds("data/model/gwRF_adaptive.rds")
```

We combine the test data with its corresponding coordinates data.

```{r}
test_data_nogeom <- cbind(
  test_data, coords_test) %>%
  st_drop_geometry()
```

#### 9.2.2 Predicting with test data

```{r}
#| eval: FALSE

gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data_nogeom, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)

GRF_pred <- write_rds(gwRF_pred, "data/model/GRF_pred.rds")
```

#### 9.2.3 Converting the predicting output into a data frame

The output of the `predict.grf()` is a vector of predicted values. It is wiser to convert it into a data frame for further visualisation and analysis.

```{r}
GRF_pred <- read_rds(
  "data/model/GRF_pred.rds")
GRF_pred_df <- as.data.frame(GRF_pred)
```

`cbind()` is used to append the predicted values onto *test_data*.

```{r}
test_data_p <- cbind(
  test_data, GRF_pred_df) %>%
  select(resale_price, GRF_pred)
```

```{r}
write_rds(test_data_p, 
          "data/model/test_data_p.rds")
```

### 9.3 Calculating Root Mean Square Error

The root mean square error (RMSE) allows us to measure how far predicted values are from observed values in a regression analysis. In the code chunk below, rmse() of Metrics package is used to compute the RMSE.

```{r}
rmse(test_data_p$resale_price, 
     test_data_p$GRF_pred)
```

### 9.4 Visualising the predicted values

Scatterplot can be used to visualise the actual resale price and the predicted resale price.

```{r}
ggplot(data = test_data_p,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```
