---
title: "8: Data Preparation for Take-home Exercise 3"
author: "Emma Chew En Chin"
date: "Oct 25, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Installing and Loading R packages

-   `httr`: Used for url call from OneMap
-   `jsonlite`: To pass the data back into tibble data format
-   `progress`: A package to track the progress of the code execution

```{r}
pacman::p_load(httr, tidyverse, sf, tmap, jsonlite, progress)
```

## 2 Geocoding HDB resale data

We can reverse geocode the HDB postal code, as there is only blk number and street name in the Resale Flats dataset.

This is the documentation: [OneMap Reverse Geocoding](https://www.onemap.gov.sg/apidocs/reverseGeocode)

```{r}
HDBresale <-read_csv("data/rawdata/HDBresale.csv") %>%
  filter(flat_type == "4 ROOM") %>%
  filter(month >= "2025-09" & month < "2025-10")
```

Changing ST to SAINT so that it does not look like 'Street'.

```{r}
HDBresale$street_name <- gsub(
  "ST\\.", "SAINT",
  HDBresale$street_name)
```

```{r}
geocode <- function(block, streetname) {
  base_url <- "https://onemap.gov.sg/api/common/elastic/search"
  address <- paste(block, streetname, sep = " ")
  query <- list("searchVal" = address,
                "returnGeom" = "Y",
                "getAddrDetails" = "N",
                "pageNum" = "1")
  
  res <-GET(base_url, query = query)
  restext <- content(res, as="text")
  
  output <- fromJSON(restext) %>%
    as.data.frame %>%z
    select(results.LATITUDE, results.LONGITUDE)
  
  return(output)
}
```

```{r}
HDBresale$LATITUDE <- 0
HDBresale$LONGITUDE <- 0

for (i in 1:nrow(HDBresale)){
  temp_output <- geocode(HDBresale[i, 4], HDBresale[i, 5])
  
  HDBresale$LATITUDE[i] <- temp_output$results.LATITUDE
  HDBresale$LONGITUDE[i] <- temp_output$results.LONGITUDE
}
```

```{r}
glimpse(HDBresale)
```

Since HDBresale contains LATITUDE and LONGITUDE columns (which are in the WGS 84 geographic coordinate system, or EPSG:4326), we need to convert it into an sf (Simple Features) object and then transform it to the SVY21 (Singapore Projected Coordinate System, EPSG:3414).

```{r}
HDBresale_sf <- HDBresale %>%
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
preschool = st_read("data/rawdata/PreSchoolsLocation.kml") %>% st_transform(crs=3414)
```

`mutate()` creates a new variable of the number of primary schools within a distance of 350m of the resale flat.

```{r}
within_350m <- st_is_within_distance(
  HDBresale_sf, preschool, dist = 350)

HDBresale_sf <- HDBresale_sf %>%
  mutate(prischool_350m = map_int(within_350m, length))
```

## 3 Preparing data for modelling

### 3.1 Installing and Loading R packages

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse,
               knitr, kableExtra, spatialRF,
               randomForestExplainer)
```

### 3.2 Data Preparation

```{r}
mdata <- read_rds("data/rawdata/mdata.rds")
```

### 3.3 Data Sampling

We will select a 10% sample at random from the data, to test as a quick prototype, as calibrating predictive models are computationally expensive.

```{r}
set.seed(1234)
HDB_sample <- mdata %>%
  sample_n(1500)
```

### 3.4 Checking of overlapping point

When using GWmodel to calibrate explanatory or predictive modesl, it is very important to ensure that there are no overlapping point features.

```{r}
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
```

### 3.5 Spatial Jitter

Using `jitter()` to move point features away from each other by 1m to avoid overlapping points.

```{r}
HDB_sample <- HDB_sample %>%
  st_jitter(amount = 1)
```

### 3.6 Data Sampling

```{r}
set.seed(1234)
resale_split <- initial_split(HDB_sample, 
                              prop = 6.67/10)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```
