---
title: "1: Geospatial Analytics for Tech HQ Location Selection"
author: "Emma Chew En Chin"
date: "Sep 11, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

Studies have shown that location is important in business competition, as business clusters often form due to the agglomeration economies effect (Porter, 1998).

In this report, we will utilise business entity location data from the Accounting and Corporate Regulatory Authority ([ACRA](https://data.gov.sg/collections/2/view)) of Singapore, to examine the location trends of clients and competitors in **Software & Applications**, **Software Publishing** and **Information Technology Consultancies**, in order to understand potential locations to situate a hypothetical Tech Consultancy Firm.

### 1.1 Data Sets

| Col1 | Col2 | Col3 |
|------|------|------|
|      |      |      |
|      |      |      |
|      |      |      |

## 2 Preparing Data for Analysis

### 2.1 Importing the R Packages

The following R packages will be used for our geospatial analysis.

```{r}
pacman::p_load(sf, tmap, spatstat, tidyverse, httr)
```

::: callout-note
-   **sf:** Used for reading, storing, manipulating, and analysing spatial data.
-   **dplyr:** From sf, used for data manipulation
-   **tmap:** Used for creating maps from spatial data.
-   **httr:** Provides a wrapper for the curl package.
-   **lubridate:** Used to handle date and date-time data type
-   **spatstat:** Used for point pattern analysis
-   **terra:** : Used to convert image output generated by spatstat into terra format
:::

### 2.2 Import ACRA Data

Data from ACRA is merged into one single dataframe.

```{r}
folder_path <- "data/aspatial/acra"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$", 
                        full.names = TRUE)
acra_data <- file_list %>%
    map_dfr(read_csv, show_col_types = FALSE)
```

```{r}
write_rds(acra_data,
          "data/rds/acra_data.rds")
```

Filtering ACRA data to companies that were registered between 2024 and 2025.

Selected business types are: - **62011:** Developing software and applications - **62021:** Information technology consultancy (except cybersecurity) - **58202:** Software publishing (software/applications) (non-games)

```{r}
my_ssic_codes <- c(62011, 62021, 58202)

biz_codes <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code %in% my_ssic_codes) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date,
                            label = TRUE,
                            abbr = TRUE)) %>%
  # if the postal code is 5 digits, add 0 in front
  filter(YEAR %in% c(2024, 2025)) %>%
  mutate(
    postal_code = str_pad(postal_code, width = 6, side = "left", pad = "0")
)
```

### 2.3 Geocoding using OneMap

Extract unique postal code and use [OneMap](https://onemap.gov.sg) to do reverse geocoding and retrieve the coordinates of the entity location.

```{r}
postcodes <- unique(biz_codes$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

# Found: geocoded, Not found: cannot be geocoded (e.g.) New development area or inaccurate postal code
found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    df <- as.data.frame(json$results, stringsAsFactors = FALSE)
    df$input_postcode <- pc 
    found <- bind_rows(found, df)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
}

print("Postal codes that were not found:")
print(not_found$postcode)
```

Remove postal codes that are not geocoded

```{r}
biz_codes_clean <- biz_codes %>%
  filter(!postal_code %in% not_found$postcode)

# Optional: Check the results
cat("Original biz_codes rows:", nrow(biz_codes), "\n")
cat("Cleaned biz_codes rows:", nrow(biz_codes_clean), "\n")
cat("Number of postal codes removed:", nrow(biz_codes) - nrow(biz_codes_clean), "\n")
```

Append location information to dataframe

```{r}
biz_codes_joined = (biz_codes_clean %>%
  left_join(found, 
            by = c('postal_code' = 'POSTAL')))
```

Save the file as an RDS file

```{r}
saveRDS(biz_codes_joined, file = "data/rds/biz_codes_joined.rds")
```

Loading the data and using X & Y coordinates as SVY21 projection

```{r}
biz_fixed <- biz_codes_joined %>%
  mutate(across(where(~is.factor(.) && nlevels(.) < 2), as.character))

biz_sf <- st_as_sf(biz_fixed,
                         coords = c("X", "Y"), 
                         crs = 3414)

class(biz_sf)
plot(st_geometry(biz_sf))
```

### 2.4 Import Map data

```{r}

```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

## References

Porter, M. E. (1998). Clusters and the new economics of competition. Harvard Business Review, 76(6), 77.
