---
title: "1: Geospatial Analytics for Tech HQ Location Selection"
author: "Emma Chew En Chin"
date: "Sep 11, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

Studies have shown that location is important in business competition, as business clusters often form due to the agglomeration economies effect (Porter, 1998).

In this report, we will utilise business entity location data from the Accounting and Corporate Regulatory Authority ([ACRA](https://data.gov.sg/collections/2/view)) of Singapore, to examine the location trends of clients and competitors in **Software & Applications**, **Software Publishing** and **Information Technology Consultancies**, in order to understand potential locations to situate a hypothetical Tech Consultancy Firm.

The primary subzones in Singapore that will be covered for this analysis is the Central Business District (CBD) consisting of subzones:

-   **Downtown Core Area** (Downtown Core, Outram)
-   **Marina Bay Precinct** (Marina South, Marina East, Singapore River)

### 1.1 Data Sets

First of all, we utilise the following geospatial data as follows:

**Geospatial Data:**

-   Master Plan 2014 Subzone Boundary (Web) from [data.gov.sg](https://data.gov.sg/)

-   LTA MRT Station Exit (KML) from [data.gov.sg](https://data.gov.sg/datasets/d_f820139ee3b0865b5512cf61ab7d1122/view)

-   ACRA Information on Corporate EntitiesAccounting from [data.gov.sg](https://data.gov.sg/collections/2/view)

### 1.2 Methodology

The objectives of this study are listed as below:

+----------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| # Objective                                                                                                                                        | # Methodology                                                                                                                                                                                                                                    |
|                                                                                                                                                    |                                                                                                                                                                                                                                                  |
| 1.  Identify the spatial distribution of points of business entities and whether there are clusters                                                | Computing **Kernel Density Estimation (KDE)** using adaptive bandwidth, to factor in unevenly distributed business entity point patterns across the CBD area..                                                                                   |
|                                                                                                                                                    |                                                                                                                                                                                                                                                  |
|                                                                                                                                                    | Discover if business entity points are randomly distributed or clustered across the CBD area., using the **L(r) function** to analyse if the points are geographically concentrated or dispersed, or displays Complete Spatial Randomness (CSR). |
+----------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 2.  Identify the spatio-temporal distribution of points of business entities, and understand seasonality trends.                                   | Visualise the geographic distribution of business entity points per month between July 2024 - June 2025 for **spatio-temporal** trend analysis.                                                                                                  |
+----------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 3.  Find out the best areas for the Tech Consultancy firm, based on distance to potential clients, competitor companies and to MRT exit locations. | Calculating a **holistic score** to identify the area within the CBD area. that is the most suitable for the Tech Consultancy Firm's office.                                                                                                     |
+----------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

## 2 Preparing Data for Analysis

### 2.1 Importing the R Packages

The following R packages will be used for our geospatial analysis.

```{r}
#| echo: false
pacman::p_load(sf, tmap, spatstat, tidyverse, httr, rvest)
```

::: callout-note
-   **sf:** Used for reading, storing, manipulating, and analysing spatial data.
-   **dplyr:** From sf, used for data manipulation
-   **tmap:** Used for creating maps from spatial data.
-   **httr:** Provides a wrapper for the curl package.
-   **lubridate:** Used to handle date and date-time data type
-   **spatstat:** Used for point pattern analysis
-   **terra:** Used to convert image output generated by spatstat into terra format
-   **rvest:** Used for scraping of data from web pages
:::

### 2.2 Import ACRA Data

Data from ACRA is merged into one single dataframe.

```{r}
#| eval: false
folder_path <- "data/aspatial/acra"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$", 
                        full.names = TRUE)
acra_data <- file_list %>%
    map_dfr(read_csv, show_col_types = FALSE)
```

```{r}
#| eval: false
write_rds(acra_data,
          "data/rds/acra_data.rds")
```

Creating separate .rds files for each business entity type, of companies that were registered between July 2024 to June 2025.

Selected business types are:

-   **62011:** Developing software and applications
-   **62021:** Information technology consultancy (except cybersecurity)
-   **58202:** Software publishing (software/applications) (non-games)

#### 2.2.1 Creating RDS file for entities under 'Developing software and applications (62011)'

```{r}
#| eval: false
biz_dev <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 62011) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date,
                            label = TRUE,
                            abbr = TRUE)) %>%
  # if the postal code is 5 digits, add 0 in front
  mutate(
    postal_code = str_pad(postal_code, 
                    
    width = 6, side = "left", pad = "0")) %>% 
  # Filter for July 2024 to June 2025
  filter((YEAR == 2024 & MONTH_NUM >= 7) | (YEAR == 2025 & MONTH_NUM <= 6))
```

Extract unique postal code and use [OneMap](https://onemap.gov.sg) to do reverse geocoding and retrieve the coordinates of the entity location.

```{r}
#| eval: false
biz_dev_postcodes <- unique(biz_dev$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

# Found: geocoded, Not found: cannot be geocoded (e.g.) New development area or inaccurate postal code
found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in biz_dev_postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    results <- bind_rows(lapply(json$results, as.data.frame))
    results$input_postcode <- pc 
    found <- bind_rows(found, results)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
  Sys.sleep(0.1)
}
```

Drop postcodes that are not geocoded, append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_dev_clean <- biz_dev %>%
  filter(!postal_code %in% not_found$postcode)

# Optional: Check the results
cat("Original biz_dev rows:", nrow(biz_dev), "\n")
cat("Cleaned biz_devs rows:", nrow(biz_dev_clean), "\n")
cat("Number of postal codes removed:", nrow(biz_dev) - nrow(biz_dev_clean), "\n")
```

Append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_dev_joined <- biz_dev_clean %>%
  left_join(found, by = c('postal_code' = 'POSTAL'), relationship = "many-to-many") %>%
  # Add jitter to avoid overlapping points
  group_by(postal_code) %>%
  mutate(
    jitter_applied = n() > 1,  # Flag if jitter was applied
    X = ifelse(jitter_applied, jitter(as.numeric(X)), as.numeric(X)),
    Y = ifelse(jitter_applied, jitter(as.numeric(Y)), as.numeric(Y))
  ) %>%
  ungroup()

saveRDS(biz_dev_joined, file = "data/rds/biz_dev_joined.rds")
```

Loading the data and using X & Y coordinates as SVY21 projection

```{r}
#| eval: false
biz_dev_fixed <- biz_dev_joined %>%
  mutate(across(where(~is.factor(.) && nlevels(.) < 2), as.character))

biz_dev_sf <- st_as_sf(biz_dev_fixed,
                         coords = c("X", "Y"), 
                         crs = 3414)

class(biz_dev_sf)
plot(st_geometry(biz_dev_sf))
```

#### 2.2.2 Creating RDS file for entities under 'Information technology consultancy (except cybersecurity)(62021)'

```{r}
#| eval: false
biz_it <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 62021) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date,
                            label = TRUE,
                            abbr = TRUE)) %>%
  # if the postal code is 5 digits, add 0 in front
  mutate(
    postal_code = str_pad(postal_code, 
                    
    width = 6, side = "left", pad = "0")) %>% 
  # Filter for July 2024 to June 2025
  filter((YEAR == 2024 & MONTH_NUM >= 7) | (YEAR == 2025 & MONTH_NUM <= 6))
```

Extract unique postal code and use [OneMap](https://onemap.gov.sg) to do reverse geocoding and retrieve the coordinates of the entity location.

```{r}
#| eval: false
biz_it_postcodes <- unique(biz_it$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

# Found: geocoded, Not found: cannot be geocoded (e.g.) New development area or inaccurate postal code
found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in biz_it_postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    results <- bind_rows(lapply(json$results, as.data.frame))
    results$input_postcode <- pc 
    found <- bind_rows(found, results)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
  Sys.sleep(0.1)
}
```

Drop postcodes that are not geocoded, append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_it_clean <- biz_it %>%
  filter(!postal_code %in% not_found$postcode)

# Optional: Check the results
cat("Original biz_it rows:", nrow(biz_it), "\n")
cat("Cleaned biz_it rows:", nrow(biz_it_clean), "\n")
cat("Number of postal codes removed:", nrow(biz_it) - nrow(biz_it_clean), "\n")
```

Append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_it_joined <- biz_it_clean %>%
  left_join(found, by = c('postal_code' = 'POSTAL'), relationship = "many-to-many") %>%
  # Add jitter to avoid overlapping points
  group_by(postal_code) %>%
  mutate(
    jitter_applied = n() > 1,  # Flag if jitter was applied
    X = ifelse(jitter_applied, jitter(as.numeric(X)), as.numeric(X)),
    Y = ifelse(jitter_applied, jitter(as.numeric(Y)), as.numeric(Y))
  ) %>%
  ungroup()

saveRDS(biz_it_joined, file = "data/rds/biz_it_joined.rds")
```

Loading the data and using X & Y coordinates as SVY21 projection

```{r}
#| eval: false
biz_it_fixed <- biz_it_joined %>%
  mutate(across(where(~is.factor(.) && nlevels(.) < 2), as.character))

biz_it_sf <- st_as_sf(biz_it_fixed,
                         coords = c("X", "Y"), 
                         crs = 3414)

class(biz_it_sf)
plot(st_geometry(biz_it_sf))
```

#### 2.2.3 Creating RDS file for entities under 'Software publishing (software/applications) (non-games) (58202)'

```{r}
#| eval: false
biz_pub <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 58202) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date,
                            label = TRUE,
                            abbr = TRUE)) %>%
  # if the postal code is 5 digits, add 0 in front
  mutate(
    postal_code = str_pad(postal_code, 
                    
    width = 6, side = "left", pad = "0")) %>% 
  # Filter for July 2024 to June 2025
  filter((YEAR == 2024 & MONTH_NUM >= 7) | (YEAR == 2025 & MONTH_NUM <= 6))
```

Extract unique postal code and use [OneMap](https://onemap.gov.sg) to do reverse geocoding and retrieve the coordinates of the entity location.

```{r}
#| eval: false
biz_pub_postcodes <- unique(biz_pub$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

# Found: geocoded, Not found: cannot be geocoded (e.g.) New development area or inaccurate postal code
found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in biz_pub_postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    results <- bind_rows(lapply(json$results, as.data.frame))
    results$input_postcode <- pc 
    found <- bind_rows(found, results)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
  Sys.sleep(0.1)
}
```

Drop postcodes that are not geocoded, append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_pub_clean <- biz_pub %>%
  filter(!postal_code %in% not_found$postcode)

# Optional: Check the results
cat("Original biz_pub rows:", nrow(biz_pub), "\n")
cat("Cleaned biz_pub rows:", nrow(biz_pub_clean), "\n")
cat("Number of postal codes removed:", nrow(biz_pub) - nrow(biz_pub_clean), "\n")
```

Append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_pub_joined <- biz_pub_clean %>%
  left_join(found, by = c('postal_code' = 'POSTAL'), relationship = "many-to-many") %>%
  # Add jitter to avoid overlapping points
  group_by(postal_code) %>%
  mutate(
    jitter_applied = n() > 1,  # Flag if jitter was applied
    X = ifelse(jitter_applied, jitter(as.numeric(X)), as.numeric(X)),
    Y = ifelse(jitter_applied, jitter(as.numeric(Y)), as.numeric(Y))
  ) %>%
  ungroup()

saveRDS(biz_pub_joined, file = "data/rds/biz_pub_joined.rds")
```

Loading the data and using X & Y coordinates as SVY21 projection

```{r}
#| eval: false
biz_pub_fixed <- biz_pub_joined %>%
  mutate(across(where(~is.factor(.) && nlevels(.) < 2), as.character))

biz_pub_sf <- st_as_sf(biz_pub_fixed,
                         coords = c("X", "Y"), 
                         crs = 3414)

class(biz_pub_sf)
plot(st_geometry(biz_pub_sf))
```

### 2.3 Import Shapefile Polygon data

```{r}
#| eval: false
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>%
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

```{r}
#| eval: false
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
#| eval: false
mpsz_sf <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
#| eval: false
mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS", 
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")

cbd_areas <- mpsz_cl %>%
  filter(PLN_AREA_N %in% c("DOWNTOWN CORE", "MARINA SOUTH", "MARINA EAST", "ROCHOR", "OUTRAM", "SINGAPORE RIVER"))
```

```{r}
#| eval: false
write_rds(mpsz_cl, 
          "data/rds/mpsz_cl.rds")
          
write_rds(cbd_areas, 
          "data/rds/cbd_areas.rds")
```

### 2.4 Import MRT location data

```{r}
#| eval: false
mrt <- st_read("data/geospatial/LTAMRTStationExitKML.kml") %>%
  st_transform(crs = 3414)
```

Save MRT data as RDS file.

```{r}
#| eval: false
saveRDS(mrt, file = "data/rds/mrt_data.rds")
```

## 3 Data Processing

### 3.1 Conversion of sf dataframes to ppp class

After importing and creating RDS files for our datasets, they will be converted to **ppp** object form.

```{r}
# Read RDS files
dev_sf <- read_rds("data/rds/biz_dev_joined.rds")
IT_sf <- read_rds("data/rds/biz_IT_joined.rds")
pub_sf <- read_rds("data/rds/biz_pub_joined.rds")
mrt_sf <- read_rds("data/rds/mrt_data.rds")
cbd_sf <- read_rds("data/rds/cbd_areas.rds")

# Convert to sf objects
dev_sf <- st_as_sf(dev_sf, coords = c("X", "Y"), crs = 3414)
IT_sf <- st_as_sf(IT_sf, coords = c("X", "Y"), crs = 3414)
pub_sf <- st_as_sf(pub_sf, coords = c("X", "Y"), crs = 3414)
mrt_sf <- st_as_sf(mrt_sf, coords = c("X", "Y"), crs = 3414)
cbd_sf <- st_as_sf(cbd_sf, coords = c("X", "Y"), crs = 3414)

# Ensure same CRS for spatial operations
cbd_sf <- st_transform(cbd_sf, st_crs(dev_sf))

# Filter business data to include only points within CBD area
dev_cbd <- dev_sf[cbd_sf, ]  # Spatial subset
IT_cbd <- IT_sf[cbd_sf, ]
pub_cbd <- pub_sf[cbd_sf, ]
mrt_cbd <- mrt_sf[cbd_sf, ]

# Convert study area into owin object
cbd_owin <- as.owin(cbd_sf)

# Apply ppp conversion
dev_ppp <- as.ppp(st_geometry(dev_cbd), W = cbd_owin)
IT_ppp <- as.ppp(st_geometry(IT_cbd), W = cbd_owin)
pub_ppp <- as.ppp(st_geometry(pub_cbd), W = cbd_owin)
mrt_ppp <- as.ppp(st_geometry(mrt_cbd), W = cbd_owin)

# Check class
class(dev_ppp)
class(IT_ppp)
class(pub_ppp)
class(mrt_ppp)
class(cbd_owin)
```

### 3.2 Combining ppp object with owin object

```{r}
dev_ppp = dev_ppp[cbd_owin]
IT_ppp = IT_ppp[cbd_owin]
pub_ppp = pub_ppp[cbd_owin]
mrt_ppp = mrt_ppp[cbd_owin]
```

## 4 Objective 1: Discover if business entity points are randomly distributed or clustered across the CBD area

1.  Identify the spatial distribution of points of business entities and whether there are clusters Computing **Kernel Density Estimation (KDE)** using adaptive bandwidth, to factor in unevenly distributed business entity point patterns across the CBD area.
2.  Discover if business entity points are randomly distributed or clustered across CBD area., using the **L(r) function** to analyse if the points are geographically concentrated or dispersed, or displays Complete Spatial Randomness (CSR).

### Spatial Point Pattern Analysis by Business Type {.tabset}

First, we employ **Kernel Density Estimation (KDE)** to visualise and analyse first-order spatial point patterns, providing insights into the intensity distribution of business entities across the CBD. To account for potential spatial heterogeneity in point patterns, we implement an adaptive bandwidth selection method using Diggle's approach (`bw.diggle`). This method applies a Gaussian kernel smoother that dynamically adjusts to local point densities, enabling more accurate identification of clustering behavior and hotspot regions within the study area.

Then, **Clark-Evan's test** is conducted to statistically confirm clustering significance, testing against complete spatial randomness (CSR) using nearest-neighbor distance analysis. This will be performed for all 3 business types at a 95% confidence interval.

The test hypotheses are: **H0**: The distribution of business entities are randomly distributed. **H1**: The distribution of business entities are not randomly distributed.

#### Software and Applications Development

##### 4.1.1 KDE for 'Developing software and applications (62011)'

As there are subtle density differences, logarithmic transformation is applied to enhance contrasting distances between points.

```{r}
#| echo: false
dev_ppp_km <- rescale.ppp(
  dev_ppp, 1000, "km")

kde_dev_diggle <- density(
  dev_ppp_km,
  sigma=bw.diggle,
  edge=TRUE,
  kernel="gaussian") 

# Apply log transformation to enhance contrasts
kde_dev_log <- kde_dev_diggle
kde_dev_log$v <- log1p(kde_dev_diggle$v)  # log(1+x) to handle zeros

# Group by Planning Area
cbd_sf_union <- cbd_sf |>
  dplyr::group_by(PLN_AREA_N) |>
  dplyr::summarise()

cbd_sf_labels <- st_point_on_surface(cbd_sf_union)
coords_km <- st_coordinates(cbd_sf_labels) / 1000

# Plot KDE + title
plot(kde_dev_log, ribbon = TRUE)
title("Software and Applications Development Companies")

# Add labels
text(coords_km,
     labels = cbd_sf_labels$PLN_AREA_N,  # change to your planning area field
     cex = 0.5, col = "white", font = 2)
```

From the plotted KDE map, we can see a major cluster in the Downtown Core - Outram area, and another more dispersed cluster at Rochor area.

##### 4.1.2 Clark-Evan's Test for 'Developing software and applications (62011)'

```{r}
#| echo: false
clarkevans.test(dev_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                method="MonteCarlo",
                nsim=99)
```

From the results, we can conclude at a 95% statistical significance level that we can reject the null hypothesis that the distribution of Software and Applications Development Companies are randomly distributed.

#### Software Publishing Companies

##### 4.1.3 KDE for 'Software publishing (software/applications) (non-games) (58202)'

```{r}
#| echo: false
pub_ppp_km <- rescale.ppp(
  pub_ppp, 1000, "km")

kde_pub_diggle <- density(
  pub_ppp_km,
  sigma=bw.diggle,
  edge=TRUE,
  kernel="gaussian") 

# Apply log transformation to enhance contrasts
kde_pub_log <- kde_pub_diggle
kde_pub_log$v <- log1p(kde_pub_diggle$v)  # log(1+x) to handle zeros

# Group by Planning Area
cbd_sf_union <- cbd_sf |>
  dplyr::group_by(PLN_AREA_N) |>
  dplyr::summarise()

cbd_sf_labels <- st_point_on_surface(cbd_sf_union)
coords_km <- st_coordinates(cbd_sf_labels) / 1000

# Plot KDE + title
plot(kde_pub_log, ribbon = TRUE)
title("Software Publishing Companies")

# Add labels
text(coords_km,
     labels = cbd_sf_labels$PLN_AREA_N,  # change to your planning area field
     cex = 0.5, col = "white", font = 2)
```

From the plotted KDE map, we can observe that smaller clusters may exist around the Downtown Core - Outram area, although they are sparse

##### 4.1.4 Clark-Evan's Test for 'Software publishing (software/applications) (non-games) (58202)'

```{r}
#| echo: false
clarkevans.test(pub_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                method="MonteCarlo",
                nsim=99)
```

#### IT Consultancy Companies

From the results, we can conclude at a 95% statistical significance level that we can reject the null hypothesis that the distribution of Software Publishing Companies are randomly distributed.

##### 4.1.5 KDE for 'Information technology consultancy (except cybersecurity)(62021)'

```{r}
#| echo: false
IT_ppp_km <- rescale.ppp(
  IT_ppp, 1000, "km")

kde_IT_diggle <- density(
  IT_ppp_km,
  sigma=bw.diggle,
  edge=TRUE,
  kernel="gaussian") 

# Apply log transformation to enhance contrasts
kde_IT_log <- kde_IT_diggle
kde_IT_log$v <- log1p(kde_IT_diggle$v)  # log(1+x) to handle zeros

# Group by Planning Area
cbd_sf_union <- cbd_sf |>
  dplyr::group_by(PLN_AREA_N) |>
  dplyr::summarise()

cbd_sf_labels <- st_point_on_surface(cbd_sf_union)
coords_km <- st_coordinates(cbd_sf_labels) / 1000

# Plot KDE + title
plot(kde_IT_log, ribbon = TRUE)
title("IT Consulting Companies")

# Add labels
text(coords_km,
     labels = cbd_sf_labels$PLN_AREA_N,  # change to your planning area field
     cex = 0.5, col = "white", font = 2)
```

From the plotted KDE map, we can see a major cluster in the Downtown Core - Outram area.

##### 4.1.6 Clark-Evan's Test for 'Information technology consultancy (except cybersecurity)(62021)'

```{r}
#| echo: false
clarkevans.test(IT_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                method="MonteCarlo",
                nsim=99)
```

From the results, we can conclude at a 95% statistical significance level that we can reject the null hypothesis that the distribution of IT Consulting Companies are randomly distributed.

### Spatial Scale Dependency

**L-function** is used to reveal how clustering behaviour varies across different spatial scales, identifying the specific distances at which clustering is most pronounced. This approach helps distinguish between localised tight clusters and broader agglomerations, providing a more nuanced understanding of the spatial organisation of business entities within the CBD area.

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

## References

Porter, M. E. (1998). Clusters and the new economics of competition. Harvard Business Review, 76(6), 77.
