---
title: "1: Geospatial Analytics for Tech HQ Location Selection"
author: "Emma Chew En Chin"
date: "Sep 11, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

Studies have shown that location is important in business competition, as business clusters often form due to the agglomeration economies effect (Porter, 1998).

In this report, we will utilise business entity location data from the Accounting and Corporate Regulatory Authority ([ACRA](https://data.gov.sg/collections/2/view)) of Singapore, to examine the location trends of clients and competitors in **Software & Applications**, **Software Publishing** and **Information Technology Consultancies**, in order to understand potential locations to situate a hypothetical Tech Consultancy Firm.

### 1.1 Data Sets

First of all, we utilise the following geospatial and aspatial data as follows:

**Geospatial Data:**

-   Master Plan 2014 Subzone Boundary (Web) from [data.gov.sg](https://data.gov.sg/)
-   LTA MRT Station Exit (KML) from [data.gov.sg](https://data.gov.sg/datasets/d_f820139ee3b0865b5512cf61ab7d1122/view)

**Aspatial Data:**

-   ACRA Information on Corporate EntitiesAccounting from [data.gov.sg](https://data.gov.sg/collections/2/view)

### 1.2 Methodology

The objectives of this study are listed as below:

+----------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| # Objective                                                                                                                                        | # Methodology                                                                                                                                                                                                                                |
|                                                                                                                                                    |                                                                                                                                                                                                                                              |
| 1.  Identify the spatial distribution of points of business entities and whether there are clusters                                                | Computing **Kernel Density Estimation (KDE)** using adaptive bandwidth, to factor in unevenly distributed business entity point patterns across Singapore.                                                                                   |
|                                                                                                                                                    |                                                                                                                                                                                                                                              |
|                                                                                                                                                    | Discover if business entity points are randomly distributed or clustered across Singapore, using the **L(r) function** to analyse if the points are geographically concentrated or dispersed, or displays Complete Spatial Randomness (CSR). |
+----------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 2.  Identify the spatio-temporal distribution of points of business entities, and understand seasonality trends.                                   | Visualise the geographic distribution of business entity points per month between July 2024 - June 2025 for **spatio-temporal** trend analysis.                                                                                              |
+----------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 3.  Find out the best areas for the Tech Consultancy firm, based on distance to potential clients, competitor companies and to MRT exit locations. | Calculating a **holistic score** to identify the area within Singapore that is the most suitable for the Tech Consultancy Firm's office.                                                                                                     |
+----------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

## 2 Preparing Data for Analysis

### 2.1 Importing the R Packages

The following R packages will be used for our geospatial analysis.

```{r}
#| echo: false
pacman::p_load(sf, tmap, spatstat, tidyverse, httr, rvest)
```

::: callout-note
-   **sf:** Used for reading, storing, manipulating, and analysing spatial data.
-   **dplyr:** From sf, used for data manipulation
-   **tmap:** Used for creating maps from spatial data.
-   **httr:** Provides a wrapper for the curl package.
-   **lubridate:** Used to handle date and date-time data type
-   **spatstat:** Used for point pattern analysis
-   **terra:** Used to convert image output generated by spatstat into terra format
-   **rvest:** Used for scraping of data from web pages
:::

### 2.2 Import ACRA Data

Data from ACRA is merged into one single dataframe.

```{r}
#| eval: false
folder_path <- "data/aspatial/acra"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$", 
                        full.names = TRUE)
acra_data <- file_list %>%
    map_dfr(read_csv, show_col_types = FALSE)
```

```{r}
#| eval: false
write_rds(acra_data,
          "data/rds/acra_data.rds")
```

Creating separate .rds files for each business entity type, of companies that were registered between July 2024 to June 2025.

Selected business types are: - **62011:** Developing software and applications - **62021:** Information technology consultancy (except cybersecurity) - **58202:** Software publishing (software/applications) (non-games)

#### 2.2.1 Creating RDS file for entities under 'Developing software and applications (62011)'

```{r}
#| eval: false
biz_dev <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 62011) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date,
                            label = TRUE,
                            abbr = TRUE)) %>%
  # if the postal code is 5 digits, add 0 in front
  mutate(
    postal_code = str_pad(postal_code, 
                    
    width = 6, side = "left", pad = "0")) %>% 
  # Filter for July 2024 to June 2025
  filter((YEAR == 2024 & MONTH_NUM >= 7) | (YEAR == 2025 & MONTH_NUM <= 6))
```

Extract unique postal code and use [OneMap](https://onemap.gov.sg) to do reverse geocoding and retrieve the coordinates of the entity location.

```{r}
#| eval: false
biz_dev_postcodes <- unique(biz_dev$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

# Found: geocoded, Not found: cannot be geocoded (e.g.) New development area or inaccurate postal code
found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in biz_dev_postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    results <- bind_rows(lapply(json$results, as.data.frame))
    results$input_postcode <- pc 
    found <- bind_rows(found, results)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
  Sys.sleep(0.1)
}
```

Drop postcodes that are not geocoded, append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_dev_clean <- biz_dev %>%
  filter(!postal_code %in% not_found$postcode)

# Optional: Check the results
cat("Original biz_dev rows:", nrow(biz_dev), "\n")
cat("Cleaned biz_devs rows:", nrow(biz_dev_clean), "\n")
cat("Number of postal codes removed:", nrow(biz_dev) - nrow(biz_dev_clean), "\n")
```

Append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_dev_joined <- biz_dev_clean %>%
  left_join(found, by = c('postal_code' = 'POSTAL'), relationship = "many-to-many") %>%
  # Add jitter to avoid overlapping points
  group_by(postal_code) %>%
  mutate(
    jitter_applied = n() > 1,  # Flag if jitter was applied
    X = ifelse(jitter_applied, jitter(as.numeric(X)), as.numeric(X)),
    Y = ifelse(jitter_applied, jitter(as.numeric(Y)), as.numeric(Y))
  ) %>%
  ungroup()

saveRDS(biz_dev_joined, file = "data/rds/biz_dev_joined.rds")
```

Loading the data and using X & Y coordinates as SVY21 projection

```{r}
#| eval: false
biz_dev_fixed <- biz_dev_joined %>%
  mutate(across(where(~is.factor(.) && nlevels(.) < 2), as.character))

biz_dev_sf <- st_as_sf(biz_dev_fixed,
                         coords = c("X", "Y"), 
                         crs = 3414)

class(biz_dev_sf)
plot(st_geometry(biz_dev_sf))
```

#### 2.2.2 Creating RDS file for entities under 'Information technology consultancy (except cybersecurity)(62021)'

```{r}
#| eval: false
biz_it <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 62021) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date,
                            label = TRUE,
                            abbr = TRUE)) %>%
  # if the postal code is 5 digits, add 0 in front
  mutate(
    postal_code = str_pad(postal_code, 
                    
    width = 6, side = "left", pad = "0")) %>% 
  # Filter for July 2024 to June 2025
  filter((YEAR == 2024 & MONTH_NUM >= 7) | (YEAR == 2025 & MONTH_NUM <= 6))
```

Extract unique postal code and use [OneMap](https://onemap.gov.sg) to do reverse geocoding and retrieve the coordinates of the entity location.

```{r}
#| eval: false
biz_it_postcodes <- unique(biz_it$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

# Found: geocoded, Not found: cannot be geocoded (e.g.) New development area or inaccurate postal code
found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in biz_it_postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    results <- bind_rows(lapply(json$results, as.data.frame))
    results$input_postcode <- pc 
    found <- bind_rows(found, results)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
  Sys.sleep(0.1)
}
```

Drop postcodes that are not geocoded, append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_it_clean <- biz_it %>%
  filter(!postal_code %in% not_found$postcode)

# Optional: Check the results
cat("Original biz_it rows:", nrow(biz_it), "\n")
cat("Cleaned biz_it rows:", nrow(biz_it_clean), "\n")
cat("Number of postal codes removed:", nrow(biz_it) - nrow(biz_it_clean), "\n")
```

Append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_it_joined <- biz_it_clean %>%
  left_join(found, by = c('postal_code' = 'POSTAL'), relationship = "many-to-many") %>%
  # Add jitter to avoid overlapping points
  group_by(postal_code) %>%
  mutate(
    jitter_applied = n() > 1,  # Flag if jitter was applied
    X = ifelse(jitter_applied, jitter(as.numeric(X)), as.numeric(X)),
    Y = ifelse(jitter_applied, jitter(as.numeric(Y)), as.numeric(Y))
  ) %>%
  ungroup()

saveRDS(biz_it_joined, file = "data/rds/biz_it_joined.rds")
```

Loading the data and using X & Y coordinates as SVY21 projection

```{r}
#| eval: false
biz_it_fixed <- biz_it_joined %>%
  mutate(across(where(~is.factor(.) && nlevels(.) < 2), as.character))

biz_it_sf <- st_as_sf(biz_it_fixed,
                         coords = c("X", "Y"), 
                         crs = 3414)

class(biz_it_sf)
plot(st_geometry(biz_it_sf))
```

#### 2.2.3 Creating RDS file for entities under 'Software publishing (software/applications) (non-games) (58202)'

```{r}
#| eval: false
biz_pub <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 58202) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date,
                            label = TRUE,
                            abbr = TRUE)) %>%
  # if the postal code is 5 digits, add 0 in front
  mutate(
    postal_code = str_pad(postal_code, 
                    
    width = 6, side = "left", pad = "0")) %>% 
  # Filter for July 2024 to June 2025
  filter((YEAR == 2024 & MONTH_NUM >= 7) | (YEAR == 2025 & MONTH_NUM <= 6))
```

Extract unique postal code and use [OneMap](https://onemap.gov.sg) to do reverse geocoding and retrieve the coordinates of the entity location.

```{r}
#| eval: false
biz_pub_postcodes <- unique(biz_pub$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

# Found: geocoded, Not found: cannot be geocoded (e.g.) New development area or inaccurate postal code
found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in biz_pub_postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    results <- bind_rows(lapply(json$results, as.data.frame))
    results$input_postcode <- pc 
    found <- bind_rows(found, results)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
  Sys.sleep(0.1)
}
```

Drop postcodes that are not geocoded, append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_pub_clean <- biz_pub %>%
  filter(!postal_code %in% not_found$postcode)

# Optional: Check the results
cat("Original biz_pub rows:", nrow(biz_pub), "\n")
cat("Cleaned biz_pub rows:", nrow(biz_pub_clean), "\n")
cat("Number of postal codes removed:", nrow(biz_pub) - nrow(biz_pub_clean), "\n")
```

Append the location data and saving the data as an RDS file.

```{r}
#| eval: false
biz_pub_joined <- biz_pub_clean %>%
  left_join(found, by = c('postal_code' = 'POSTAL'), relationship = "many-to-many") %>%
  # Add jitter to avoid overlapping points
  group_by(postal_code) %>%
  mutate(
    jitter_applied = n() > 1,  # Flag if jitter was applied
    X = ifelse(jitter_applied, jitter(as.numeric(X)), as.numeric(X)),
    Y = ifelse(jitter_applied, jitter(as.numeric(Y)), as.numeric(Y))
  ) %>%
  ungroup()

saveRDS(biz_pub_joined, file = "data/rds/biz_pub_joined.rds")
```

Loading the data and using X & Y coordinates as SVY21 projection

```{r}
#| eval: false
biz_pub_fixed <- biz_pub_joined %>%
  mutate(across(where(~is.factor(.) && nlevels(.) < 2), as.character))

biz_pub_sf <- st_as_sf(biz_pub_fixed,
                         coords = c("X", "Y"), 
                         crs = 3414)

class(biz_pub_sf)
plot(st_geometry(biz_pub_sf))
```

### 2.3 Import Shapefile Polygon data

```{r}
#| eval: false
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>%
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

```{r}
#| eval: false
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
#| eval: false
mpsz_sf <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
#| eval: false
mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS",
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")
```

```{r}
#| eval: false
write_rds(mpsz_cl, 
          "data/mpsz_cl.rds")
```

### 2.4 Import MRT location data

```{r}
#| eval: false
mrt <- st_read("data/geospatial/LTAMRTStationExitKML.kml") %>%
  st_transform(crs = 3414)
```

```{r}
#| eval: false
plot(st_geometry(mpsz_cl), col = "grey")
plot(st_geometry(mrt), add = TRUE, pch = 19, col = "black", cex = 0.6)
```

Save MRT data as RDS file.

```{r}
#| eval: false
saveRDS(mrt, file = "data/rds/mrt_data.rds")
```

## 3 


### Read biz_sf dataframe

```{r}
biz_sf <- read_rds("data/rds/biz_dev_joined.rds")
```

## References

Porter, M. E. (1998). Clusters and the new economics of competition. Harvard Business Review, 76(6), 77.
