---
title: "Hands-on Exercise 1"
author: "Emma Chew En Chin"
date: "Aug 20, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## Getting Started

Install and launch R packages

The code chunk below uses p_load of pacman package to check if tidyverse package have been installed.

```{r}
pacman::p_load(sf, tidyverse, here)
```

## Importing the Data

Importing polygon feature data in shapefile format

```{r}
mpsz = st_read(dsn = "data/geospatial/MasterPlan2014SubzoneBoundaryWebSHP", 
               layer = "MP14_SUBZONE_WEB_PL")
```
Importing polyline feature data in shapefile format

```{r}
cyclingpath = st_read(dsn = "data/geospatial/CyclingPath_Apr2025", 
                         layer = "CyclingPathGazette")
```
Importing GIS data in kml format
```{r}
preschool = st_read("data/geospatial/PreSchoolsLocation.kml")
```
## Data Exploration
st_geometry() displays basic geospatial information.
```{r}
st_geometry(mpsz)
```
glimpse() shows the associated attribute information in the dataset.
```{r}
glimpse(mpsz)
```
head() shows complete information of a feature object.
```{r}
head(mpsz, n=5)  
```
## Plotting Geospatial Data
Below is a multi-plot of all attributes.
```{r}
plot(mpsz)
```
Only the geometry can be plotted as well.
```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

```{r}
head(mpsz, n=5)  
```