---
title: "4: Using GWmodel using sp dataframe"
author: "Emma Chew En Chin"
date: "Sep 20, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

Using GWmodel using sp dataframe

## 2 Loading the Data

### 2.1 Importing packages into R environment

```{r}
pacman::p_load(sf, ggstatsplot, tmap, tidyverse, knitr, GWmodel)
```

### 2.2 Importing shapefile into R environment

```{r}
hunan_sf <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### 2.3 Importing csv into R environment

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

Left Join Hunan and Hunan 2012 dataset

```{r}
hunan_sf <- left_join(hunan_sf, hunan2012) %>%
  select(1:3, 7, 15, 16, 21, 32)
```

Convert to sp

```{r}
hunan_sp <- hunan_sf %>%
  as_Spatial()
```

### 3 Geographically Weighted Summary Statistics

:::: panel_tabset
::: {style="font-size: 1.20em"}
### Cross Validation

```{r}
bw_CV <- bw.gwr(GDPPC ~ 1,
                data = hunan_sp,
                approach = "CV",
                adaptive = FALSE,
                kernel = "bisquare",
                longlat = T)
```

### AIC

```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1,
                data = hunan_sp,
                approach = "AIC",
                adaptive = FALSE,
                kernel = "bisquare",
                longlat = T)
```

### GWSS

```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               kernel = "bisquare",
               adaptive = FALSE,
              longlat = T)
```
:::

### 3.1 Extract SDF Data table from GWSS object output

```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

`cbind()` joins two tables with same row sequence without an ID (join) column.

```{r}
hunan_gstat <- cbind(hunan_sf, gwstat_df)
```

