---
title: "4: Using GWmodel using sp dataframe"
author: "Emma Chew En Chin"
date: "Sep 20, 2025"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 1 Overview

Using GWmodel using sp dataframe

## 2 Loading the Data

### 2.1 Importing packages into R environment

```{r}
pacman::p_load(sf, ggstatsplot, tmap, tidyverse, knitr, GWmodel)
```

### 2.2 Importing shapefile into R environment

```{r}
hunan_sf <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### 2.3 Importing csv into R environment

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

Left Join Hunan and Hunan 2012 dataset

```{r}
hunan_sf <- left_join(hunan_sf, hunan2012) %>%
  select(1:3, 7, 15, 16, 21, 32)
```

Convert to sp

```{r}
hunan_sp <- hunan_sf %>%
  as_Spatial()
```

### 3 Mapping GDPPC

```{r}
basemap <- tm_shape(hunan_sf) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan_sf, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)

hunan_sp <- hunan_sf %>%
  as_Spatial()
```

### 3 Geographically Weighted Summary Statistics

:::: panel_tabset
::: {style="font-size: 1.20em"}
### Cross Validation

```{r}
bw_CV <- bw.gwr(GDPPC ~ 1, 
             data = hunan_sp,
             approach = "CV",
             adaptive = TRUE, 
             kernel = "bisquare", 
             longlat = T)
```

### AIC

```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1, 
             data = hunan_sp,
             approach ="AIC",
             adaptive = TRUE, 
             kernel = "bisquare", 
             longlat = T)
```

### GWSS

```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               kernel = "bisquare",
               quantile = TRUE,
               adaptive = TRUE,
               longlat = T)
```
:::

### 3.1 Extract SDF Data table from GWSS object output

```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

`cbind()` joins two tables with same row sequence without an ID (join) column.

```{r}
hunan_gstat <- cbind(hunan_sf, gwstat_df)
```
::::

### 3.2 Visualising Geographically Weighted Summary Statistics

```{r}
tm_shape(hunan_gstat) +
  tm_polygons(fill = "GDPPC_LM",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "local Mean")) +
  tm_borders(fill_alpha = 0.5) +
  tm_layout(main.title = "Distribution of geographically wieghted mean",
            main.title.position = "center",
            main.title.size = 2.0,
            frame = TRUE)
```


### 3.3 Geographically Weighted Correlation

Business question: Is there any relationship between GDP per capita and Gross Industry Output?

```{r}
ggscatterstats(
  data = hunan2012, 
  x = Agri, 
  y = GDPPC,
  xlab = "Gross Agriculture Output", ## label for the x-axis
  ylab = "GDP per capita", 
  label.var = County, 
  label.expression = Agri > 10000 & GDPPC > 50000, 
  point.label.args = list(alpha = 0.7, size = 4, color = "grey50"),
  xfill = "#CC79A7", 
  yfill = "#009E73", 
  title = "Relationship between GDP PC and Gross Agriculture Output")
```

## 4 Geospatial analytics solution

```{r}
bw <- bw.gwr(GDPPC ~ GIO, 
             data = hunan_sp, 
             approach = "AICc", 
             adaptive = TRUE)
```

```{r}
gwstats <- gwss(hunan_sp, 
                vars = c("GDPPC", "GIO"), 
                bw = bw,
                kernel = "bisquare",
                adaptive = TRUE, 
                longlat = T)
```

```{r}
gwstat_df <- as.data.frame(gwstats$SDF) %>%
  select(c(12,13)) %>%
  rename(gwCorr = Corr_GDPPC.GIO,
         gwSpearman = Spearman_rho_GDPPC.GIO)
```

```{r}
hunan_Corr <- cbind(hunan_sf, gwstat_df)
```

## 5 Visualising Local Correlation

```{r}
tm_shape(hunan_Corr) +
  tm_polygons(fill = "gwSpearman",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "local Spearman Rho")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Local Spearman Rho", 
  size = 2.0) +
  tm_layout(frame = TRUE)
```